### 2023-03-17 doc/multiverse.Rmd
Add analysis where study is a fixed effect instead of a random effect. unnest participants from within measurement days. Nest measurement days within studies by use of concatenate. Create rmd for multiverse.

---
title             : "Multiverse analysis"
shorttitle        : "Title"

author: 
  - name          : "First Author"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "my@email.com"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - "Conceptualization"
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"
  - name          : "Ernst-August Doelle"
    affiliation   : "1,2"
    role:
      - "Writing - Review & Editing"
      - "Supervision"

affiliation:
  - id            : "1"
    institution   : "Wilhelm-Wundt-University"
  - id            : "2"
    institution   : "Konstanz Business School"

authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.

abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.

  <!-- https://tinyurl.com/ybremelq -->

keywords          : "keywords"
wordcount         : "X"

floatsintext      : no
linenumbers       : no
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no

classoption       : "man"
output            : papaja::apa6_docx
---

\renewcommand{\arraystretch}{0.65}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
source("R/utils.R")
options(warn = -1)
knitr::opts_chunk$set(out.width = "110%")
```

```{r load-libraries}
library(targets)
library(gt)
require(ggplot2)
library(papaja)
```

```{r load-targets, include=FALSE}
model_list_by_age_fixedef <- tar_read(model_list_by_age_fixedef)
tar_load(data_clean)
tar_load(purdy_pictures_by_age_fixedef)
model_tables <- tar_read(model_tables_fixedef)
pretty <- function(x) format(x, big.mark = ",")
```

# Results

```{r}
n_participants <- length(unique(data_clean$participant_id))
```

## Multiverse

In this document we show the results using another analytic technique. Namely treating study ID as a fixed effect. This is what we wrote in the protocol originally.

## The effects of physical activity volume on sleep

We estimated the effects of physical activity on sleep (RQ1) using mixed-effects models
We estimated the effect of physical activity volume on sleep by age, and the results are presented in Table \@ref(tab:sleep-outcomes) and Figure \@ref(fig:sleep-by-volume-fig)

```{r sleep-outcomes}
papaja::apa_table(
  model_tables$sleep$data,
  caption = model_tables$sleep$caption,
  note = model_tables$sleep$note,
  col_spanners = model_tables$sleep$col_spanners,
  escape = FALSE,
  landscape = TRUE
)
```

```{r sleep-by-volume-fig, fig.cap = "Sleep metrics on Physical activity volume"}
knitr::include_graphics(here::here("Figures/sleep on pa_volume by Age_fixedef.jpg"), error = FALSE)
```

## The effects of physical activity intensity on sleep

We estimated how physical activity intensity affects sleep across different age groups
We present the results controlling for sex, SES, and BMI, in Table \@ref(tab:sleep-outcomes) and Figure \@ref(fig:sleep-by-intensity-fig).

```{r sleep-by-intensity-fig, fig.cap = "Sleep metrics on Physical activity intensity"}
knitr::include_graphics(here::here("Figures/sleep on pa_intensity by Age_fixedef.jpg"), error = FALSE)
```

## The effects of sleep duration on physical activity

We estimated the effect of sleep duration on physical activity by age
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes) and Figure \@ref(fig:PA-by-sleep-duration-fig).

```{r pa-outcomes}
papaja::apa_table(
  model_tables$physical_activity$data,
  caption = model_tables$physical_activity$caption,
  note = model_tables$physical_activity$note,
  col_spanners = model_tables$physical_activity$col_spanners,
  escape = FALSE,
  landscape = TRUE
)
```

```{r PA-by-sleep-duration-fig, fig.cap = "Physical activty by sleep duration"}
knitr::include_graphics(here::here("Figures/PA on sleep_duration_lag by Age_fixedef.jpg"), error = FALSE)
```

## The effects of sleep efficiency on physical activity

We estimated the effect of sleep efficiency on physical activity by age.
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes) and Figure \@ref(fig:PA-by-sleep-efficiency-fig).

```{r PA-by-sleep-efficiency-fig, fig.cap = "Physical activty by sleep efficiency"}
knitr::include_graphics(here::here("Figures/PA on sleep_efficiency_lag by Age_fixedef.jpg"), error = FALSE)
```

## The effects of sleep onset on physical activity

We estimated the effect of sleep onset on physical activity by age
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes) and Figure \@ref(fig:PA-by-sleep-onset-fig).

```{r PA-by-sleep-onset-fig, fig.cap = "Physical activty by sleep onset"}
knitr::include_graphics(here::here("Figures/PA on sleep_onset_lag by Age_fixedef.jpg"), error = FALSE)
```

## The effects of sleep regularity on physical activity

We estimated the effect of sleep regularity on physical activity by age.
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes) and Figure \@ref(fig:PA-by-sleep-regularity-fig).

```{r PA-by-sleep-regularity-fig, fig.cap = "Physical activty by sleep regularity"}
knitr::include_graphics(here::here("Figures/PA on sleep_regularity_lag by Age_fixedef.jpg"), error = FALSE)
```