### 2023-03-21 doc/multiverse.Rmd

---
title             : "Multiverse analysis"
shorttitle        : "Title"

author: 
  - name          : "First Author"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "my@email.com"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - "Conceptualization"
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"
  - name          : "Ernst-August Doelle"
    affiliation   : "1,2"
    role:
      - "Writing - Review & Editing"
      - "Supervision"

affiliation:
  - id            : "1"
    institution   : "Wilhelm-Wundt-University"
  - id            : "2"
    institution   : "Konstanz Business School"

authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.

abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.

  <!-- https://tinyurl.com/ybremelq -->

keywords          : "keywords"
wordcount         : "X"

floatsintext      : no
linenumbers       : no
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no

classoption       : "man"

header-includes:
 - \usepackage[maxfloats=256]{morefloats}

output            : papaja::apa6_pdf

---

```{comment}
I needed to add maxdeadcycles and more floats to support how many figures are in this doc.
```

\renewcommand{\arraystretch}{0.7}
\maxdeadcycles=1000

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.pos = "!htbp")
source("R/utils.R")
options(warn = -1)
knitr::opts_chunk$set(out.width = "110%")
```

```{r load-libraries}
library(targets)
library(gt)
require(ggplot2)
library(papaja)
```

```{r load-targets, include=FALSE}
model_list_by_age_fixedef <- tar_read(model_list_by_age_fixedef)
tar_load(data_clean)
tar_load(purdy_pictures_by_age_fixedef)
model_tables <- tar_read(model_tables_fixedef)
pretty <- function(x) format(x, big.mark = ",")
```

# Results

```{r}
n_participants <- length(unique(data_clean$participant_id))
```

## Multiverse

In this document we show the results using another analytic technique. Namely treating study ID as a fixed effect. This is what we wrote in the protocol originally.

## The effects of physical activity volume on sleep

We estimated the effects of physical activity on sleep (RQ1) using mixed-effects models
We estimated the effect of physical activity volume on sleep by age, and the results are presented in Table \@ref(tab:sleep-outcomes) and Figure \@ref(fig:sleep-by-volume-fig)

```{r sleep-outcomes}
papaja::apa_table(
  model_tables$sleep$data,
  caption = model_tables$sleep$caption,
  note = model_tables$sleep$note,
  col_spanners = model_tables$sleep$col_spanners,
  escape = FALSE,
  landscape = TRUE
)
```

```{r sleep-by-volume-fig, fig.cap = "Sleep metrics on Physical activity volume"}
knitr::include_graphics(here::here("Figures/sleep on pa_volume by Age_fixedef.jpg"), error = FALSE)
```

## The effects of physical activity intensity on sleep

We estimated how physical activity intensity affects sleep across different age groups
We present the results controlling for sex, SES, and BMI, in Table \@ref(tab:sleep-outcomes) and Figure \@ref(fig:sleep-by-intensity-fig).

```{r sleep-by-intensity-fig, fig.cap = "Sleep metrics on Physical activity intensity"}
knitr::include_graphics(here::here("Figures/sleep on pa_intensity by Age_fixedef.jpg"), error = FALSE)
```

## The effects of sleep duration on physical activity

We estimated the effect of sleep duration on physical activity by age
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes) and Figure \@ref(fig:PA-by-sleep-duration-fig).

```{r pa-outcomes}
papaja::apa_table(
  model_tables$physical_activity$data,
  caption = model_tables$physical_activity$caption,
  note = model_tables$physical_activity$note,
  col_spanners = model_tables$physical_activity$col_spanners,
  escape = FALSE,
  landscape = TRUE
)
```

```{r PA-by-sleep-duration-fig, fig.cap = "Physical activity by sleep duration"}
knitr::include_graphics(here::here("Figures/PA on sleep_duration_lag by Age_fixedef.jpg"), error = FALSE)
```

## The effects of sleep efficiency on physical activity

We estimated the effect of sleep efficiency on physical activity by age.
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes) and Figure \@ref(fig:PA-by-sleep-efficiency-fig).

```{r PA-by-sleep-efficiency-fig, fig.cap = "Physical activity by sleep efficiency"}
knitr::include_graphics(here::here("Figures/PA on sleep_efficiency_lag by Age_fixedef.jpg"), error = FALSE)
```

## The effects of sleep onset on physical activity

We estimated the effect of sleep onset on physical activity by age
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes) and Figure \@ref(fig:PA-by-sleep-onset-fig).

```{r PA-by-sleep-onset-fig, fig.cap = "Physical activity by sleep onset"}
knitr::include_graphics(here::here("Figures/PA on sleep_onset_lag by Age_fixedef.jpg"), error = FALSE)
```

## The effects of sleep regularity on physical activity

We estimated the effect of sleep regularity on physical activity by age.
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes) and Figure \@ref(fig:PA-by-sleep-regularity-fig).

```{r PA-by-sleep-regularity-fig, fig.cap = "Physical activity by sleep regularity"}
knitr::include_graphics(here::here("Figures/PA on sleep_regularity_lag by Age_fixedef.jpg"), error = FALSE)
```

# Moderation

In the following section we present models for a range of moderators which were pre-defined in our protocol.

```{r eval = FALSE}
pattern <- readLines(fs::path("doc/results_chunk.md")) |> paste(collapse = "\n")

moderator_df <- data.frame(
  moderator = c("bmi", "ses", "sex", "weekday", "season", "region", "daylight", "wear_location"),
  moderator_formal = c("BMI", "SES", "sex", "weekday", "season", "region", "daylight hours", "wear location"),
  image_path = c("Bmi", "Ses", "Sex", "Weekday_x", "Season", "Region", "Daylight_hours", "Accelerometer_wear_location")
)
moderator_df$moderator_sentence <- stringr::str_to_sentence(moderator_df$moderator)

lapply(seq_len(nrow(moderator_df)), function(r){

  moderator <- moderator_df$moderator[r]
  moderator_fig <- gsub("_","-",moderator)
  moderator_formal <- moderator_df$moderator_formal[r]
  moderator_sentence <- moderator_df$moderator_sentence[r]
  image_path <- moderator_df$image_path[r]

  glue::glue(pattern, .open = "_::", .close = "::_")

}) |> paste(collapse = "\n") |> clipr::write_clip()

```

<!------------------------------------>
<!-- Auto-generated chunks go here: -->
<!------------------------------------>

\clearpage

## BMI

```{r}
tar_load(model_tables_bmi)
```

### The effects of physical activity volume on sleep

We estimated the effects of physical activity on sleep (RQ1) using mixed-effects models
We estimated the effect of physical activity volume on sleep by BMI, and the results are presented in Table \@ref(tab:sleep-outcomes-by-bmi) and Figure \@ref(fig:sleep-by-volume-fig-by-bmi)

```{r sleep-outcomes-by-bmi}
papaja::apa_table(
  model_tables_bmi$sleep$data,
  caption = model_tables_bmi$sleep$caption,
  note = model_tables_bmi$sleep$note,
  col_spanners = model_tables_bmi$sleep$col_spanners,
  escape = FALSE,
  landscape = TRUE
)
```

```{r sleep-by-volume-fig-by-bmi, fig.cap = "Sleep metrics on Physical activity volume by BMI"}
knitr::include_graphics(here::here("Figures/Sleep on pa_volume by Bmi.jpg"), error = FALSE)
```

### The effects of physical activity intensity on sleep by BMI

We estimated how physical activity intensity affects sleep across BMI
We present the results in Table \@ref(tab:sleep-outcomes-by-bmi) and Figure \@ref(fig:sleep-by-intensity-fig-by-bmi).

```{r sleep-by-intensity-fig-by-bmi, fig.cap = "Sleep metrics on Physical activity intensity moderated by BMI"}
knitr::include_graphics(here::here("Figures/Sleep on pa_intensity by Bmi.jpg"), error = FALSE)
```

### The effects of sleep duration on physical activity

We estimated the effect of sleep duration on physical activity by BMI
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes) and Figure \@ref(fig:PA-by-sleep-duration-fig-by-bmi).

```{r pa-outcomes-by-bmi}
papaja::apa_table(
  model_tables_bmi$physical_activity$data,
  caption = model_tables_bmi$physical_activity$caption,
  note = model_tables_bmi$physical_activity$note,
  col_spanners = model_tables_bmi$physical_activity$col_spanners,
  escape = FALSE,
  landscape = TRUE
)
```

```{r PA-by-sleep-duration-fig-by-bmi, fig.cap = "Physical activity by sleep duration moderated by BMI"}
knitr::include_graphics(here::here("Figures/Pa on sleep_duration_lag by Bmi.jpg"), error = FALSE)
```

### The effects of sleep efficiency on physical activity

We estimated the effect of sleep efficiency on physical activity by BMI.
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes-by-bmi) and Figure \@ref(fig:PA-by-sleep-efficiency-fig-by-bmi).

```{r PA-by-sleep-efficiency-fig-by-bmi, fig.cap = "Physical activity by sleep efficiency moderated by BMI"}
knitr::include_graphics(here::here("Figures/Pa on sleep_efficiency_lag by Bmi.jpg"), error = FALSE)
```

### The effects of sleep onset on physical activity

We estimated the effect of sleep onset on physical activity by BMI
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes) and Figure \@ref(fig:PA-by-sleep-onset-fig-by-bmi).

```{r PA-by-sleep-onset-fig-by-bmi, fig.cap = "Physical activity by sleep onset moderated by BMI"}
knitr::include_graphics(here::here("Figures/Pa on sleep_onset_lag by Bmi.jpg"), error = FALSE)
```

### The effects of sleep regularity on physical activity

We estimated the effect of sleep regularity on physical activity by BMI.
Results are presented in Table \@ref(tab:pa-outcomes) and Figure \@ref(fig:PA-by-sleep-regularity-fig-by-bmi).

```{r PA-by-sleep-regularity-fig-by-bmi, fig.cap = "Physical activity by sleep regularity moderated by BMI"}
knitr::include_graphics(here::here("Figures/Pa on sleep_regularity_lag by Bmi.jpg"), error = FALSE)
```
\clearpage

## SES

```{r}
tar_load(model_tables_ses)
```

### The effects of physical activity volume on sleep

We estimated the effects of physical activity on sleep (RQ1) using mixed-effects models
We estimated the effect of physical activity volume on sleep by SES, and the results are presented in Table \@ref(tab:sleep-outcomes-by-ses) and Figure \@ref(fig:sleep-by-volume-fig-by-ses)

```{r sleep-outcomes-by-ses}
papaja::apa_table(
  model_tables_ses$sleep$data,
  caption = model_tables_ses$sleep$caption,
  note = model_tables_ses$sleep$note,
  col_spanners = model_tables_ses$sleep$col_spanners,
  escape = FALSE,
  landscape = TRUE
)
```

```{r sleep-by-volume-fig-by-ses, fig.cap = "Sleep metrics on Physical activity volume by SES"}
knitr::include_graphics(here::here("Figures/Sleep on pa_volume by Ses.jpg"), error = FALSE)
```

### The effects of physical activity intensity on sleep by SES

We estimated how physical activity intensity affects sleep across SES
We present the results in Table \@ref(tab:sleep-outcomes-by-ses) and Figure \@ref(fig:sleep-by-intensity-fig-by-ses).

```{r sleep-by-intensity-fig-by-ses, fig.cap = "Sleep metrics on Physical activity intensity moderated by SES"}
knitr::include_graphics(here::here("Figures/Sleep on pa_intensity by Ses.jpg"), error = FALSE)
```

### The effects of sleep duration on physical activity

We estimated the effect of sleep duration on physical activity by SES
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes) and Figure \@ref(fig:PA-by-sleep-duration-fig-by-ses).

```{r pa-outcomes-by-ses}
papaja::apa_table(
  model_tables_ses$physical_activity$data,
  caption = model_tables_ses$physical_activity$caption,
  note = model_tables_ses$physical_activity$note,
  col_spanners = model_tables_ses$physical_activity$col_spanners,
  escape = FALSE,
  landscape = TRUE
)
```

```{r PA-by-sleep-duration-fig-by-ses, fig.cap = "Physical activity by sleep duration moderated by SES"}
knitr::include_graphics(here::here("Figures/Pa on sleep_duration_lag by Ses.jpg"), error = FALSE)
```

### The effects of sleep efficiency on physical activity

We estimated the effect of sleep efficiency on physical activity by SES.
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes-by-ses) and Figure \@ref(fig:PA-by-sleep-efficiency-fig-by-ses).

```{r PA-by-sleep-efficiency-fig-by-ses, fig.cap = "Physical activity by sleep efficiency moderated by SES"}
knitr::include_graphics(here::here("Figures/Pa on sleep_efficiency_lag by Ses.jpg"), error = FALSE)
```

### The effects of sleep onset on physical activity

We estimated the effect of sleep onset on physical activity by SES
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes) and Figure \@ref(fig:PA-by-sleep-onset-fig-by-ses).

```{r PA-by-sleep-onset-fig-by-ses, fig.cap = "Physical activity by sleep onset moderated by SES"}
knitr::include_graphics(here::here("Figures/Pa on sleep_onset_lag by Ses.jpg"), error = FALSE)
```

### The effects of sleep regularity on physical activity

We estimated the effect of sleep regularity on physical activity by SES.
Results are presented in Table \@ref(tab:pa-outcomes) and Figure \@ref(fig:PA-by-sleep-regularity-fig-by-ses).

```{r PA-by-sleep-regularity-fig-by-ses, fig.cap = "Physical activity by sleep regularity moderated by SES"}
knitr::include_graphics(here::here("Figures/Pa on sleep_regularity_lag by Ses.jpg"), error = FALSE)
```
\clearpage

## sex

```{r}
tar_load(model_tables_sex)
```

### The effects of physical activity volume on sleep

We estimated the effects of physical activity on sleep (RQ1) using mixed-effects models
We estimated the effect of physical activity volume on sleep by sex, and the results are presented in Table \@ref(tab:sleep-outcomes-by-sex) and Figure \@ref(fig:sleep-by-volume-fig-by-sex)

```{r sleep-outcomes-by-sex}
papaja::apa_table(
  model_tables_sex$sleep$data,
  caption = model_tables_sex$sleep$caption,
  note = model_tables_sex$sleep$note,
  col_spanners = model_tables_sex$sleep$col_spanners,
  escape = FALSE,
  landscape = TRUE
)
```

```{r sleep-by-volume-fig-by-sex, fig.cap = "Sleep metrics on Physical activity volume by sex"}
knitr::include_graphics(here::here("Figures/Sleep on pa_volume by Sex.jpg"), error = FALSE)
```

### The effects of physical activity intensity on sleep by sex

We estimated how physical activity intensity affects sleep across sex
We present the results in Table \@ref(tab:sleep-outcomes-by-sex) and Figure \@ref(fig:sleep-by-intensity-fig-by-sex).

```{r sleep-by-intensity-fig-by-sex, fig.cap = "Sleep metrics on Physical activity intensity moderated by sex"}
knitr::include_graphics(here::here("Figures/Sleep on pa_intensity by Sex.jpg"), error = FALSE)
```

### The effects of sleep duration on physical activity

We estimated the effect of sleep duration on physical activity by sex
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes) and Figure \@ref(fig:PA-by-sleep-duration-fig-by-sex).

```{r pa-outcomes-by-sex}
papaja::apa_table(
  model_tables_sex$physical_activity$data,
  caption = model_tables_sex$physical_activity$caption,
  note = model_tables_sex$physical_activity$note,
  col_spanners = model_tables_sex$physical_activity$col_spanners,
  escape = FALSE,
  landscape = TRUE
)
```

```{r PA-by-sleep-duration-fig-by-sex, fig.cap = "Physical activity by sleep duration moderated by sex"}
knitr::include_graphics(here::here("Figures/Pa on sleep_duration_lag by Sex.jpg"), error = FALSE)
```

### The effects of sleep efficiency on physical activity

We estimated the effect of sleep efficiency on physical activity by sex.
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes-by-sex) and Figure \@ref(fig:PA-by-sleep-efficiency-fig-by-sex).

```{r PA-by-sleep-efficiency-fig-by-sex, fig.cap = "Physical activity by sleep efficiency moderated by sex"}
knitr::include_graphics(here::here("Figures/Pa on sleep_efficiency_lag by Sex.jpg"), error = FALSE)
```

### The effects of sleep onset on physical activity

We estimated the effect of sleep onset on physical activity by sex
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes) and Figure \@ref(fig:PA-by-sleep-onset-fig-by-sex).

```{r PA-by-sleep-onset-fig-by-sex, fig.cap = "Physical activity by sleep onset moderated by sex"}
knitr::include_graphics(here::here("Figures/Pa on sleep_onset_lag by Sex.jpg"), error = FALSE)
```

### The effects of sleep regularity on physical activity

We estimated the effect of sleep regularity on physical activity by sex.
Results are presented in Table \@ref(tab:pa-outcomes) and Figure \@ref(fig:PA-by-sleep-regularity-fig-by-sex).

```{r PA-by-sleep-regularity-fig-by-sex, fig.cap = "Physical activity by sleep regularity moderated by sex"}
knitr::include_graphics(here::here("Figures/Pa on sleep_regularity_lag by Sex.jpg"), error = FALSE)
```
\clearpage

## weekday

```{r}
tar_load(model_tables_weekday)
```

### The effects of physical activity volume on sleep

We estimated the effects of physical activity on sleep (RQ1) using mixed-effects models
We estimated the effect of physical activity volume on sleep by weekday, and the results are presented in Table \@ref(tab:sleep-outcomes-by-weekday) and Figure \@ref(fig:sleep-by-volume-fig-by-weekday)

```{r sleep-outcomes-by-weekday}
papaja::apa_table(
  model_tables_weekday$sleep$data,
  caption = model_tables_weekday$sleep$caption,
  note = model_tables_weekday$sleep$note,
  col_spanners = model_tables_weekday$sleep$col_spanners,
  escape = FALSE,
  landscape = TRUE
)
```

```{r sleep-by-volume-fig-by-weekday, fig.cap = "Sleep metrics on Physical activity volume by weekday"}
knitr::include_graphics(here::here("Figures/Sleep on pa_volume by Weekday_x.jpg"), error = FALSE)
```

### The effects of physical activity intensity on sleep by weekday

We estimated how physical activity intensity affects sleep across weekday
We present the results in Table \@ref(tab:sleep-outcomes-by-weekday) and Figure \@ref(fig:sleep-by-intensity-fig-by-weekday).

```{r sleep-by-intensity-fig-by-weekday, fig.cap = "Sleep metrics on Physical activity intensity moderated by weekday"}
knitr::include_graphics(here::here("Figures/Sleep on pa_intensity by Weekday_x.jpg"), error = FALSE)
```

### The effects of sleep duration on physical activity

We estimated the effect of sleep duration on physical activity by weekday
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes) and Figure \@ref(fig:PA-by-sleep-duration-fig-by-weekday).

```{r pa-outcomes-by-weekday}
papaja::apa_table(
  model_tables_weekday$physical_activity$data,
  caption = model_tables_weekday$physical_activity$caption,
  note = model_tables_weekday$physical_activity$note,
  col_spanners = model_tables_weekday$physical_activity$col_spanners,
  escape = FALSE,
  landscape = TRUE
)
```

```{r PA-by-sleep-duration-fig-by-weekday, fig.cap = "Physical activity by sleep duration moderated by weekday"}
knitr::include_graphics(here::here("Figures/Pa on sleep_duration_lag by Weekday_x.jpg"), error = FALSE)
```

### The effects of sleep efficiency on physical activity

We estimated the effect of sleep efficiency on physical activity by weekday.
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes-by-weekday) and Figure \@ref(fig:PA-by-sleep-efficiency-fig-by-weekday).

```{r PA-by-sleep-efficiency-fig-by-weekday, fig.cap = "Physical activity by sleep efficiency moderated by weekday"}
knitr::include_graphics(here::here("Figures/Pa on sleep_efficiency_lag by Weekday_x.jpg"), error = FALSE)
```

### The effects of sleep onset on physical activity

We estimated the effect of sleep onset on physical activity by weekday
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes) and Figure \@ref(fig:PA-by-sleep-onset-fig-by-weekday).

```{r PA-by-sleep-onset-fig-by-weekday, fig.cap = "Physical activity by sleep onset moderated by weekday"}
knitr::include_graphics(here::here("Figures/Pa on sleep_onset_lag by Weekday_x.jpg"), error = FALSE)
```

### The effects of sleep regularity on physical activity

We estimated the effect of sleep regularity on physical activity by weekday.
Results are presented in Table \@ref(tab:pa-outcomes) and Figure \@ref(fig:PA-by-sleep-regularity-fig-by-weekday).

```{r PA-by-sleep-regularity-fig-by-weekday, fig.cap = "Physical activity by sleep regularity moderated by weekday"}
knitr::include_graphics(here::here("Figures/Pa on sleep_regularity_lag by Weekday_x.jpg"), error = FALSE)
```
\clearpage

## season

```{r}
tar_load(model_tables_season)
```

### The effects of physical activity volume on sleep

We estimated the effects of physical activity on sleep (RQ1) using mixed-effects models
We estimated the effect of physical activity volume on sleep by season, and the results are presented in Table \@ref(tab:sleep-outcomes-by-season) and Figure \@ref(fig:sleep-by-volume-fig-by-season)

```{r sleep-outcomes-by-season}
papaja::apa_table(
  model_tables_season$sleep$data,
  caption = model_tables_season$sleep$caption,
  note = model_tables_season$sleep$note,
  col_spanners = model_tables_season$sleep$col_spanners,
  escape = FALSE,
  landscape = TRUE
)
```

```{r sleep-by-volume-fig-by-season, fig.cap = "Sleep metrics on Physical activity volume by season"}
knitr::include_graphics(here::here("Figures/Sleep on pa_volume by Season.jpg"), error = FALSE)
```

### The effects of physical activity intensity on sleep by season

We estimated how physical activity intensity affects sleep across season
We present the results in Table \@ref(tab:sleep-outcomes-by-season) and Figure \@ref(fig:sleep-by-intensity-fig-by-season).

```{r sleep-by-intensity-fig-by-season, fig.cap = "Sleep metrics on Physical activity intensity moderated by season"}
knitr::include_graphics(here::here("Figures/Sleep on pa_intensity by Season.jpg"), error = FALSE)
```

### The effects of sleep duration on physical activity

We estimated the effect of sleep duration on physical activity by season
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes) and Figure \@ref(fig:PA-by-sleep-duration-fig-by-season).

```{r pa-outcomes-by-season}
papaja::apa_table(
  model_tables_season$physical_activity$data,
  caption = model_tables_season$physical_activity$caption,
  note = model_tables_season$physical_activity$note,
  col_spanners = model_tables_season$physical_activity$col_spanners,
  escape = FALSE,
  landscape = TRUE
)
```

```{r PA-by-sleep-duration-fig-by-season, fig.cap = "Physical activity by sleep duration moderated by season"}
knitr::include_graphics(here::here("Figures/Pa on sleep_duration_lag by Season.jpg"), error = FALSE)
```

### The effects of sleep efficiency on physical activity

We estimated the effect of sleep efficiency on physical activity by season.
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes-by-season) and Figure \@ref(fig:PA-by-sleep-efficiency-fig-by-season).

```{r PA-by-sleep-efficiency-fig-by-season, fig.cap = "Physical activity by sleep efficiency moderated by season"}
knitr::include_graphics(here::here("Figures/Pa on sleep_efficiency_lag by Season.jpg"), error = FALSE)
```

### The effects of sleep onset on physical activity

We estimated the effect of sleep onset on physical activity by season
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes) and Figure \@ref(fig:PA-by-sleep-onset-fig-by-season).

```{r PA-by-sleep-onset-fig-by-season, fig.cap = "Physical activity by sleep onset moderated by season"}
knitr::include_graphics(here::here("Figures/Pa on sleep_onset_lag by Season.jpg"), error = FALSE)
```

### The effects of sleep regularity on physical activity

We estimated the effect of sleep regularity on physical activity by season.
Results are presented in Table \@ref(tab:pa-outcomes) and Figure \@ref(fig:PA-by-sleep-regularity-fig-by-season).

```{r PA-by-sleep-regularity-fig-by-season, fig.cap = "Physical activity by sleep regularity moderated by season"}
knitr::include_graphics(here::here("Figures/Pa on sleep_regularity_lag by Season.jpg"), error = FALSE)
```
\clearpage

## region

```{r}
tar_load(model_tables_region)
```

### The effects of physical activity volume on sleep

We estimated the effects of physical activity on sleep (RQ1) using mixed-effects models
We estimated the effect of physical activity volume on sleep by region, and the results are presented in Table \@ref(tab:sleep-outcomes-by-region) and Figure \@ref(fig:sleep-by-volume-fig-by-region)

```{r sleep-outcomes-by-region}
papaja::apa_table(
  model_tables_region$sleep$data,
  caption = model_tables_region$sleep$caption,
  note = model_tables_region$sleep$note,
  col_spanners = model_tables_region$sleep$col_spanners,
  escape = FALSE,
  landscape = TRUE
)
```

```{r sleep-by-volume-fig-by-region, fig.cap = "Sleep metrics on Physical activity volume by region"}
knitr::include_graphics(here::here("Figures/Sleep on pa_volume by Region.jpg"), error = FALSE)
```

### The effects of physical activity intensity on sleep by region

We estimated how physical activity intensity affects sleep across region
We present the results in Table \@ref(tab:sleep-outcomes-by-region) and Figure \@ref(fig:sleep-by-intensity-fig-by-region).

```{r sleep-by-intensity-fig-by-region, fig.cap = "Sleep metrics on Physical activity intensity moderated by region"}
knitr::include_graphics(here::here("Figures/Sleep on pa_intensity by Region.jpg"), error = FALSE)
```

### The effects of sleep duration on physical activity

We estimated the effect of sleep duration on physical activity by region
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes) and Figure \@ref(fig:PA-by-sleep-duration-fig-by-region).

```{r pa-outcomes-by-region}
papaja::apa_table(
  model_tables_region$physical_activity$data,
  caption = model_tables_region$physical_activity$caption,
  note = model_tables_region$physical_activity$note,
  col_spanners = model_tables_region$physical_activity$col_spanners,
  escape = FALSE,
  landscape = TRUE
)
```

```{r PA-by-sleep-duration-fig-by-region, fig.cap = "Physical activity by sleep duration moderated by region"}
knitr::include_graphics(here::here("Figures/Pa on sleep_duration_lag by Region.jpg"), error = FALSE)
```

### The effects of sleep efficiency on physical activity

We estimated the effect of sleep efficiency on physical activity by region.
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes-by-region) and Figure \@ref(fig:PA-by-sleep-efficiency-fig-by-region).

```{r PA-by-sleep-efficiency-fig-by-region, fig.cap = "Physical activity by sleep efficiency moderated by region"}
knitr::include_graphics(here::here("Figures/Pa on sleep_efficiency_lag by Region.jpg"), error = FALSE)
```

### The effects of sleep onset on physical activity

We estimated the effect of sleep onset on physical activity by region
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes) and Figure \@ref(fig:PA-by-sleep-onset-fig-by-region).

```{r PA-by-sleep-onset-fig-by-region, fig.cap = "Physical activity by sleep onset moderated by region"}
knitr::include_graphics(here::here("Figures/Pa on sleep_onset_lag by Region.jpg"), error = FALSE)
```

### The effects of sleep regularity on physical activity

We estimated the effect of sleep regularity on physical activity by region.
Results are presented in Table \@ref(tab:pa-outcomes) and Figure \@ref(fig:PA-by-sleep-regularity-fig-by-region).

```{r PA-by-sleep-regularity-fig-by-region, fig.cap = "Physical activity by sleep regularity moderated by region"}
knitr::include_graphics(here::here("Figures/Pa on sleep_regularity_lag by Region.jpg"), error = FALSE)
```
\clearpage

## daylight hours

```{r}
tar_load(model_tables_daylight)
```

### The effects of physical activity volume on sleep

We estimated the effects of physical activity on sleep (RQ1) using mixed-effects models
We estimated the effect of physical activity volume on sleep by daylight hours, and the results are presented in Table \@ref(tab:sleep-outcomes-by-daylight) and Figure \@ref(fig:sleep-by-volume-fig-by-daylight)

```{r sleep-outcomes-by-daylight}
papaja::apa_table(
  model_tables_daylight$sleep$data,
  caption = model_tables_daylight$sleep$caption,
  note = model_tables_daylight$sleep$note,
  col_spanners = model_tables_daylight$sleep$col_spanners,
  escape = FALSE,
  landscape = TRUE
)
```

```{r sleep-by-volume-fig-by-daylight, fig.cap = "Sleep metrics on Physical activity volume by daylight hours"}
knitr::include_graphics(here::here("Figures/Sleep on pa_volume by Daylight_hours.jpg"), error = FALSE)
```

### The effects of physical activity intensity on sleep by daylight hours

We estimated how physical activity intensity affects sleep across daylight hours
We present the results in Table \@ref(tab:sleep-outcomes-by-daylight) and Figure \@ref(fig:sleep-by-intensity-fig-by-daylight).

```{r sleep-by-intensity-fig-by-daylight, fig.cap = "Sleep metrics on Physical activity intensity moderated by daylight hours"}
knitr::include_graphics(here::here("Figures/Sleep on pa_intensity by Daylight_hours.jpg"), error = FALSE)
```

### The effects of sleep duration on physical activity

We estimated the effect of sleep duration on physical activity by daylight hours
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes) and Figure \@ref(fig:PA-by-sleep-duration-fig-by-daylight).

```{r pa-outcomes-by-daylight}
papaja::apa_table(
  model_tables_daylight$physical_activity$data,
  caption = model_tables_daylight$physical_activity$caption,
  note = model_tables_daylight$physical_activity$note,
  col_spanners = model_tables_daylight$physical_activity$col_spanners,
  escape = FALSE,
  landscape = TRUE
)
```

```{r PA-by-sleep-duration-fig-by-daylight, fig.cap = "Physical activity by sleep duration moderated by daylight hours"}
knitr::include_graphics(here::here("Figures/Pa on sleep_duration_lag by Daylight_hours.jpg"), error = FALSE)
```

### The effects of sleep efficiency on physical activity

We estimated the effect of sleep efficiency on physical activity by daylight hours.
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes-by-daylight) and Figure \@ref(fig:PA-by-sleep-efficiency-fig-by-daylight).

```{r PA-by-sleep-efficiency-fig-by-daylight, fig.cap = "Physical activity by sleep efficiency moderated by daylight hours"}
knitr::include_graphics(here::here("Figures/Pa on sleep_efficiency_lag by Daylight_hours.jpg"), error = FALSE)
```

### The effects of sleep onset on physical activity

We estimated the effect of sleep onset on physical activity by daylight hours
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes) and Figure \@ref(fig:PA-by-sleep-onset-fig-by-daylight).

```{r PA-by-sleep-onset-fig-by-daylight, fig.cap = "Physical activity by sleep onset moderated by daylight hours"}
knitr::include_graphics(here::here("Figures/Pa on sleep_onset_lag by Daylight_hours.jpg"), error = FALSE)
```

### The effects of sleep regularity on physical activity

We estimated the effect of sleep regularity on physical activity by daylight hours.
Results are presented in Table \@ref(tab:pa-outcomes) and Figure \@ref(fig:PA-by-sleep-regularity-fig-by-daylight).

```{r PA-by-sleep-regularity-fig-by-daylight, fig.cap = "Physical activity by sleep regularity moderated by daylight hours"}
knitr::include_graphics(here::here("Figures/Pa on sleep_regularity_lag by Daylight_hours.jpg"), error = FALSE)
```
\clearpage

## wear location

```{r}
tar_load(model_tables_wear_location)
```

### The effects of physical activity volume on sleep

We estimated the effects of physical activity on sleep (RQ1) using mixed-effects models
We estimated the effect of physical activity volume on sleep by wear location, and the results are presented in Table \@ref(tab:sleep-outcomes-by-wear_location) and Figure \@ref(fig:sleep-by-volume-fig-by-wear-location)

```{r sleep-outcomes-by-wear_location}
papaja::apa_table(
  model_tables_wear_location$sleep$data,
  caption = model_tables_wear_location$sleep$caption,
  note = model_tables_wear_location$sleep$note,
  col_spanners = model_tables_wear_location$sleep$col_spanners,
  escape = FALSE,
  landscape = TRUE
)
```

```{r sleep-by-volume-fig-by-wear-location, fig.cap = "Sleep metrics on Physical activity volume by wear location"}
knitr::include_graphics(here::here("Figures/Sleep on pa_volume by Accelerometer_wear_location.jpg"), error = FALSE)
```

### The effects of physical activity intensity on sleep by wear location

We estimated how physical activity intensity affects sleep across wear location
We present the results in Table \@ref(tab:sleep-outcomes-by-wear_location) and Figure \@ref(fig:sleep-by-intensity-fig-by-wear-location).

```{r sleep-by-intensity-fig-by-wear-location, fig.cap = "Sleep metrics on Physical activity intensity moderated by wear location"}
knitr::include_graphics(here::here("Figures/Sleep on pa_intensity by Accelerometer_wear_location.jpg"), error = FALSE)
```

### The effects of sleep duration on physical activity

We estimated the effect of sleep duration on physical activity by wear location
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes) and Figure \@ref(fig:PA-by-sleep-duration-fig-by-wear-location).

```{r pa-outcomes-by-wear_location}
papaja::apa_table(
  model_tables_wear_location$physical_activity$data,
  caption = model_tables_wear_location$physical_activity$caption,
  note = model_tables_wear_location$physical_activity$note,
  col_spanners = model_tables_wear_location$physical_activity$col_spanners,
  escape = FALSE,
  landscape = TRUE
)
```

```{r PA-by-sleep-duration-fig-by-wear-location, fig.cap = "Physical activity by sleep duration moderated by wear location"}
knitr::include_graphics(here::here("Figures/Pa on sleep_duration_lag by Accelerometer_wear_location.jpg"), error = FALSE)
```

### The effects of sleep efficiency on physical activity

We estimated the effect of sleep efficiency on physical activity by wear location.
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes-by-wear_location) and Figure \@ref(fig:PA-by-sleep-efficiency-fig-by-wear-location).

```{r PA-by-sleep-efficiency-fig-by-wear-location, fig.cap = "Physical activity by sleep efficiency moderated by wear location"}
knitr::include_graphics(here::here("Figures/Pa on sleep_efficiency_lag by Accelerometer_wear_location.jpg"), error = FALSE)
```

### The effects of sleep onset on physical activity

We estimated the effect of sleep onset on physical activity by wear location
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes) and Figure \@ref(fig:PA-by-sleep-onset-fig-by-wear-location).

```{r PA-by-sleep-onset-fig-by-wear-location, fig.cap = "Physical activity by sleep onset moderated by wear location"}
knitr::include_graphics(here::here("Figures/Pa on sleep_onset_lag by Accelerometer_wear_location.jpg"), error = FALSE)
```

### The effects of sleep regularity on physical activity

We estimated the effect of sleep regularity on physical activity by wear location.
Results are presented in Table \@ref(tab:pa-outcomes) and Figure \@ref(fig:PA-by-sleep-regularity-fig-by-wear-location).

```{r PA-by-sleep-regularity-fig-by-wear-location, fig.cap = "Physical activity by sleep regularity moderated by wear location"}
knitr::include_graphics(here::here("Figures/Pa on sleep_regularity_lag by Accelerometer_wear_location.jpg"), error = FALSE)
```

<!-- Finish auto-generated chunks go here: -->