### 2023-04-04 doc/multiverse_data_imp_log.Rmd

---
title             : "Multiverse analysis"
shorttitle        : "Title"

author: 
  - name          : "First Author"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "my@email.com"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - "Conceptualization"
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"
  - name          : "Ernst-August Doelle"
    affiliation   : "1,2"
    role:
      - "Writing - Review & Editing"
      - "Supervision"

affiliation:
  - id            : "1"
    institution   : "Wilhelm-Wundt-University"
  - id            : "2"
    institution   : "Konstanz Business School"

authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.

abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.

  <!-- https://tinyurl.com/ybremelq -->

keywords          : "keywords"
wordcount         : "X"

floatsintext      : no
linenumbers       : no
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no

classoption       : "man"

header-includes:
 - \usepackage[maxfloats=256]{morefloats}

output            : papaja::apa6_pdf

---

```{comment}
I needed to add maxdeadcycles and more floats to support how many figures are in this doc.
```

\renewcommand{\arraystretch}{0.7}
\maxdeadcycles=1000

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.pos = "!htbp")
source("R/utils.R")
options(warn = -1)
knitr::opts_chunk$set(out.width = "110%")
```

```{r load-libraries}
library(targets)
library(gt)
require(ggplot2)
library(papaja)
```

```{r load-targets, include=FALSE}
tar_load(data_clean)
pretty <- function(x) format(x, big.mark = ",")
```

# Results

```{r}
n_participants <- length(unique(data_clean$participant_id))
```

## Model diagnostics

The model diagnostics for the main analysis, and for all models presented in this document are in Table \@ref(tab:model-diagnostics)

```{r model-diagnostics}
tar_read(model_diagnostics) |> 
  papaja::apa_table(caption = "Model diagnostics", escape = FALSE,
                    longtable = TRUE)
```

## Multiverse

In this document we show the relationship between sleep and physical activity by the moderators that were pre-registered in our protocol.

# Moderation

In the following section we present models for a range of moderators which were pre-defined in our protocol.

```{r eval = FALSE}
pattern <- readLines(fs::path("doc/results_chunk.md")) |> paste(collapse = "\n")

moderator_df <- data.frame(
  moderator = c("bmi", "ses", "sex", "weekday_x", "season", "region", "daylight_hours", "accelerometer_wear_location", "pa_mostactivehr"),
  moderator_formal = c("BMI", "SES", "sex", "weekday", "season", "region", "daylight hours", "wear location", "most active hour"),
  image_path = c("Bmi", "Ses", "Sex", "Weekday_x", "Season", "Region", "Daylight_hours", "Accelerometer_wear_location", "Pa_mostactivehr")
)

lapply(seq_len(nrow(moderator_df)), function(r){

  moderator <- moderator_df$moderator[r]
  moderator_fig <- gsub("_","-",moderator)
  moderator_formal <- moderator_df$moderator_formal[r]
  image_path <- moderator_df$image_path[r]

  glue::glue(pattern, .open = "_::", .close = "::_")

}) |> paste(collapse = "\n") |> clipr::write_clip()

```

<!------------------------------------>
<!-- Auto-generated chunks go here: -->
<!------------------------------------>

<!-- Finish auto-generated chunks go here: -->