### 2023-04-13 doc/multiverse.Rmd

---
title             : "Multiverse analysis"
shorttitle        : "Title"

author: 
  - name          : "First Author"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "my@email.com"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - "Conceptualization"
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"
  - name          : "Ernst-August Doelle"
    affiliation   : "1,2"
    role:
      - "Writing - Review & Editing"
      - "Supervision"

affiliation:
  - id            : "1"
    institution   : "Wilhelm-Wundt-University"
  - id            : "2"
    institution   : "Konstanz Business School"

authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.

abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.

  <!-- https://tinyurl.com/ybremelq -->

keywords          : "keywords"
wordcount         : "X"

floatsintext      : no
linenumbers       : no
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no

classoption       : "man"

header-includes:
 - \usepackage[maxfloats=256]{morefloats}

output            : papaja::apa6_pdf

---

```{comment}
I needed to add maxdeadcycles and more floats to support how many figures are in this doc.
```

\renewcommand{\arraystretch}{0.7}
\maxdeadcycles=1000

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.pos = "!htbp")
source("R/utils.R")
options(warn = -1)
knitr::opts_chunk$set(out.width = "110%")
```

```{r load-libraries}
library(targets)
library(gt)
require(ggplot2)
library(papaja)
```

```{r load-targets, include=FALSE}
tar_load(multiverse_file)
pretty <- function(x) format(x, big.mark = ",")
```

# Results

## Model diagnostics

The model diagnostics for the main analysis, and for all models presented in this document are in Table \@ref(tab:model-diagnostics)

```{r model-diagnostics}
tar_read(model_diagnostics) |>
  papaja::apa_table(
    caption = "Model diagnostics", escape = FALSE,
    longtable = TRUE
  )
```

## Fixed Effect Model

```{r}
tar_load(model_tables_by_age_fixedef)
tar_load(purdy_pictures_by_age_fixedef)
```

In this document we show the results using another analytic technique. 
Namely treating study ID as a fixed effect. 
This is what we wrote in the protocol originally.

### The effects of physical activity volume on sleep

We estimated the effects of physical activity on sleep (RQ1) using fixed-effects models.
We estimated the effect of physical activity volume on sleep by age, and the results are presented in Table \@ref(tab:sleep-outcomes-fixedef) and Figure \@ref(fig:sleep-by-volume-fig-fixedef).

```{r sleep-outcomes-fixedef}
papaja::apa_table(
  model_tables_by_age_fixedef$sleep$data,
  caption = model_tables_by_age_fixedef$sleep$caption,
  note = model_tables_by_age_fixedef$sleep$note,
  col_spanners = model_tables_by_age_fixedef$sleep$col_spanners,
  escape = FALSE,
  landscape = TRUE
)
```

```{r sleep-by-volume-fig-fixedef, fig.cap = "Sleep metrics on Physical activity volume"}
knitr::include_graphics(here::here(purdy_pictures_by_age_fixedef$predictor_pa_volume), error = FALSE)
```

### The effects of physical activity intensity on sleep

We estimated how physical activity intensity affects sleep across different age groups.
We present the results controlling for sex, SES, and BMI, in Table \@ref(tab:sleep-outcomes-fixedef) and Figure \@ref(fig:sleep-by-intensity-fig-fixedef).

```{r sleep-by-intensity-fig-fixedef, fig.cap = "Sleep metrics on Physical activity intensity"}
knitr::include_graphics(here::here(purdy_pictures_by_age_fixedef$predictor_pa_intensity), error = FALSE)
```

### The effects of sleep duration on physical activity

We estimated the effect of sleep duration on physical activity by age.
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes-fixedef) and Figure \@ref(fig:PA-by-sleep-duration-fig-fixedef).

```{r pa-outcomes-fixedef}
papaja::apa_table(
  model_tables_by_age_fixedef$physical_activity$data,
  caption = model_tables_by_age_fixedef$physical_activity$caption,
  note = model_tables_by_age_fixedef$physical_activity$note,
  col_spanners = model_tables_by_age_fixedef$physical_activity$col_spanners,
  escape = FALSE,
  landscape = TRUE
)
```

```{r PA-by-sleep-duration-fig-fixedef, fig.cap = "Physical activity by sleep duration"}
knitr::include_graphics(here::here(purdy_pictures_by_age_fixedef$predictor_sleep_duration_lag), error = FALSE)
```

### The effects of sleep efficiency on physical activity

We estimated the effect of sleep efficiency on physical activity by age.
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes-fixedef) and Figure \@ref(fig:PA-by-sleep-efficiency-fig-fixedef).

```{r PA-by-sleep-efficiency-fig-fixedef, fig.cap = "Physical activity by sleep efficiency"}
knitr::include_graphics(here::here(purdy_pictures_by_age_fixedef$predictor_sleep_efficiency_lag), error = FALSE)
```

### The effects of sleep onset on physical activity

We estimated the effect of sleep onset on physical activity by age.
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes-fixedef) and Figure \@ref(fig:PA-by-sleep-onset-fig-fixedef).

```{r PA-by-sleep-onset-fig-fixedef, fig.cap = "Physical activity by sleep onset"}
knitr::include_graphics(here::here(purdy_pictures_by_age_fixedef$predictor_sleep_onset_lag), error = FALSE)
```

### The effects of sleep regularity on physical activity

We estimated the effect of sleep regularity on physical activity by age.
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes-fixedef) and Figure \@ref(fig:PA-by-sleep-regularity-fig-fixedef).

```{r PA-by-sleep-regularity-fig-fixedef, fig.cap = "Physical activity by sleep regularity"}
knitr::include_graphics(here::here(purdy_pictures_by_age_fixedef$predictor_sleep_regularity_lag), error = FALSE)
```

## Without Log Transforming PA Volume

```{r}
tar_load(model_tables_by_age_nolog)
tar_load(purdy_pictures_by_age_nolog)
```

Based on preliminary modelling on a subset of the data, we chose to log transform the physical activity volume variable.
Below, we present the results had we not made this decision.
These results are all moderated by age, as this was our primary analysis.

### The effects of physical activity volume on sleep

We estimated the effects of physical activity on sleep (RQ1) using mixed-effects models.
We estimated the effect of physical activity volume on sleep by age, and the results are presented in Table \@ref(tab:sleep-outcomes-nolog) and Figure \@ref(fig:sleep-by-volume-fig-nolog).

```{r sleep-outcomes-nolog}
papaja::apa_table(
  model_tables_by_age_nolog$sleep$data,
  caption = model_tables_by_age_nolog$sleep$caption,
  note = model_tables_by_age_nolog$sleep$note,
  col_spanners = model_tables_by_age_nolog$sleep$col_spanners,
  escape = FALSE,
  landscape = TRUE
)
```

```{r sleep-by-volume-fig-nolog, fig.cap = "Sleep metrics on Physical activity volume"}
knitr::include_graphics(here::here(purdy_pictures_by_age_nolog$predictor_pa_volume), error = FALSE)
```

### The effects of sleep duration on physical activity

We estimated the effect of sleep duration on physical activity by age.
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes-nolog) and Figure \@ref(fig:PA-by-sleep-duration-fig-nolog).

```{r pa-outcomes-nolog}
papaja::apa_table(
  model_tables_by_age_nolog$physical_activity$data,
  caption = model_tables_by_age_nolog$physical_activity$caption,
  note = model_tables_by_age_nolog$physical_activity$note,
  col_spanners = model_tables_by_age_nolog$physical_activity$col_spanners,
  escape = FALSE,
  landscape = TRUE
)
```

```{r PA-by-sleep-duration-fig-nolog, fig.cap = "Physical activity by sleep duration"}
knitr::include_graphics(here::here(purdy_pictures_by_age_nolog$predictor_sleep_duration_lag), error = FALSE)
```

### The effects of sleep efficiency on physical activity

We estimated the effect of sleep efficiency on physical activity by age.
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes-nolog) and Figure \@ref(fig:PA-by-sleep-efficiency-fig-nolog).

```{r PA-by-sleep-efficiency-fig-nolog, fig.cap = "Physical activity by sleep efficiency"}
knitr::include_graphics(here::here(purdy_pictures_by_age_nolog$predictor_sleep_efficiency_lag), error = FALSE)
```

### The effects of sleep onset on physical activity

We estimated the effect of sleep onset on physical activity by age.
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes-nolog) and Figure \@ref(fig:PA-by-sleep-onset-fig-nolog).

```{r PA-by-sleep-onset-fig-nolog, fig.cap = "Physical activity by sleep onset"}
knitr::include_graphics(here::here(purdy_pictures_by_age_nolog$predictor_sleep_onset_lag), error = FALSE)
```

### The effects of sleep regularity on physical activity

We estimated the effect of sleep regularity on physical activity by age.
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes-nolog) and Figure \@ref(fig:PA-by-sleep-regularity-fig-nolog).

```{r PA-by-sleep-regularity-fig-nolog, fig.cap = "Physical activity by sleep regularity"}
knitr::include_graphics(here::here(purdy_pictures_by_age_nolog$predictor_sleep_regularity_lag), error = FALSE)
```

# Moderation

In the following section we present models for a range of moderators which were pre-defined in our protocol.
\clearpage

## BMI

```{r}
tar_load(model_tables_by_bmi)
tar_load(purdy_pictures_by_bmi)
```

### The effects of physical activity volume on sleep by BMI

We estimated the effects of physical activity on sleep (RQ1) using mixed-effects models.
We estimated the effect of physical activity volume on sleep by BMI, and the results are presented in Table \@ref(tab:sleep-outcomes-by-bmi) and Figure \@ref(fig:sleep-by-volume-fig-by-bmi)

```{r sleep-outcomes-by-bmi}
papaja::apa_table(
  model_tables_by_bmi$sleep$data,
  caption = model_tables_by_bmi$sleep$caption,
  note = model_tables_by_bmi$sleep$note,
  col_spanners = model_tables_by_bmi$sleep$col_spanners,
  escape = FALSE,
  landscape = TRUE
)
```

```{r sleep-by-volume-fig-by-bmi, fig.cap = "Sleep metrics on Physical activity volume by BMI"}
knitr::include_graphics(here::here(purdy_pictures_by_bmi$predictor_pa_volume), error = FALSE)
```

### The effects of physical activity intensity on sleep by BMI

We estimated how physical activity intensity affects sleep across BMI.
We present the results in Table \@ref(tab:sleep-outcomes-by-bmi) and Figure \@ref(fig:sleep-by-intensity-fig-by-bmi).

```{r sleep-by-intensity-fig-by-bmi, fig.cap = "Sleep metrics on Physical activity intensity moderated by BMI"}
knitr::include_graphics(here::here(purdy_pictures_by_bmi$predictor_pa_intensity), error = FALSE)
```

### The effects of sleep duration on physical activity

We estimated the effect of sleep duration on physical activity by BMI.
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes-by-bmi) and Figure \@ref(fig:PA-by-sleep-duration-fig-by-bmi).

```{r pa-outcomes-by-bmi}
papaja::apa_table(
  model_tables_by_bmi$physical_activity$data,
  caption = model_tables_by_bmi$physical_activity$caption,
  note = model_tables_by_bmi$physical_activity$note,
  col_spanners = model_tables_by_bmi$physical_activity$col_spanners,
  escape = FALSE,
  landscape = TRUE
)
```

```{r PA-by-sleep-duration-fig-by-bmi, fig.cap = "Physical activity by sleep duration moderated by BMI"}
knitr::include_graphics(here::here(purdy_pictures_by_bmi$predictor_sleep_duration_lag), error = FALSE)
```

### The effects of sleep efficiency on physical activity

We estimated the effect of sleep efficiency on physical activity by BMI.
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes-by-bmi) and Figure \@ref(fig:PA-by-sleep-efficiency-fig-by-bmi).

```{r PA-by-sleep-efficiency-fig-by-bmi, fig.cap = "Physical activity by sleep efficiency moderated by BMI"}
knitr::include_graphics(here::here(purdy_pictures_by_bmi$predictor_sleep_efficiency_lag), error = FALSE)
```

### The effects of sleep onset on physical activity

We estimated the effect of sleep onset on physical activity by BMI.
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes-by-bmi) and Figure \@ref(fig:PA-by-sleep-onset-fig-by-bmi).

```{r PA-by-sleep-onset-fig-by-bmi, fig.cap = "Physical activity by sleep onset moderated by BMI"}
knitr::include_graphics(here::here(purdy_pictures_by_bmi$predictor_sleep_onset_lag), error = FALSE)
```

### The effects of sleep regularity on physical activity

We estimated the effect of sleep regularity on physical activity by BMI.
Results are presented in Table \@ref(tab:pa-outcomes-by-bmi) and Figure \@ref(fig:PA-by-sleep-regularity-fig-by-bmi).

```{r PA-by-sleep-regularity-fig-by-bmi, fig.cap = "Physical activity by sleep regularity moderated by BMI"}
knitr::include_graphics(here::here(purdy_pictures_by_bmi$predictor_sleep_regularity_lag), error = FALSE)
```
\clearpage

## SES

```{r}
tar_load(model_tables_by_ses)
tar_load(purdy_pictures_by_ses)
```

### The effects of physical activity volume on sleep by SES

We estimated the effects of physical activity on sleep (RQ1) using mixed-effects models.
We estimated the effect of physical activity volume on sleep by SES, and the results are presented in Table \@ref(tab:sleep-outcomes-by-ses) and Figure \@ref(fig:sleep-by-volume-fig-by-ses)

```{r sleep-outcomes-by-ses}
papaja::apa_table(
  model_tables_by_ses$sleep$data,
  caption = model_tables_by_ses$sleep$caption,
  note = model_tables_by_ses$sleep$note,
  col_spanners = model_tables_by_ses$sleep$col_spanners,
  escape = FALSE,
  landscape = TRUE
)
```

```{r sleep-by-volume-fig-by-ses, fig.cap = "Sleep metrics on Physical activity volume by SES"}
knitr::include_graphics(here::here(purdy_pictures_by_ses$predictor_pa_volume), error = FALSE)
```

### The effects of physical activity intensity on sleep by SES

We estimated how physical activity intensity affects sleep across SES.
We present the results in Table \@ref(tab:sleep-outcomes-by-ses) and Figure \@ref(fig:sleep-by-intensity-fig-by-ses).

```{r sleep-by-intensity-fig-by-ses, fig.cap = "Sleep metrics on Physical activity intensity moderated by SES"}
knitr::include_graphics(here::here(purdy_pictures_by_ses$predictor_pa_intensity), error = FALSE)
```

### The effects of sleep duration on physical activity

We estimated the effect of sleep duration on physical activity by SES.
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes-by-ses) and Figure \@ref(fig:PA-by-sleep-duration-fig-by-ses).

```{r pa-outcomes-by-ses}
papaja::apa_table(
  model_tables_by_ses$physical_activity$data,
  caption = model_tables_by_ses$physical_activity$caption,
  note = model_tables_by_ses$physical_activity$note,
  col_spanners = model_tables_by_ses$physical_activity$col_spanners,
  escape = FALSE,
  landscape = TRUE
)
```

```{r PA-by-sleep-duration-fig-by-ses, fig.cap = "Physical activity by sleep duration moderated by SES"}
knitr::include_graphics(here::here(purdy_pictures_by_ses$predictor_sleep_duration_lag), error = FALSE)
```

### The effects of sleep efficiency on physical activity

We estimated the effect of sleep efficiency on physical activity by SES.
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes-by-ses) and Figure \@ref(fig:PA-by-sleep-efficiency-fig-by-ses).

```{r PA-by-sleep-efficiency-fig-by-ses, fig.cap = "Physical activity by sleep efficiency moderated by SES"}
knitr::include_graphics(here::here(purdy_pictures_by_ses$predictor_sleep_efficiency_lag), error = FALSE)
```

### The effects of sleep onset on physical activity

We estimated the effect of sleep onset on physical activity by SES.
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes-by-ses) and Figure \@ref(fig:PA-by-sleep-onset-fig-by-ses).

```{r PA-by-sleep-onset-fig-by-ses, fig.cap = "Physical activity by sleep onset moderated by SES"}
knitr::include_graphics(here::here(purdy_pictures_by_ses$predictor_sleep_onset_lag), error = FALSE)
```

### The effects of sleep regularity on physical activity

We estimated the effect of sleep regularity on physical activity by SES.
Results are presented in Table \@ref(tab:pa-outcomes-by-ses) and Figure \@ref(fig:PA-by-sleep-regularity-fig-by-ses).

```{r PA-by-sleep-regularity-fig-by-ses, fig.cap = "Physical activity by sleep regularity moderated by SES"}
knitr::include_graphics(here::here(purdy_pictures_by_ses$predictor_sleep_regularity_lag), error = FALSE)
```
\clearpage

## sex

```{r}
tar_load(model_tables_by_sex)
tar_load(purdy_pictures_by_sex)
```

### The effects of physical activity volume on sleep by sex

We estimated the effects of physical activity on sleep (RQ1) using mixed-effects models.
We estimated the effect of physical activity volume on sleep by sex, and the results are presented in Table \@ref(tab:sleep-outcomes-by-sex) and Figure \@ref(fig:sleep-by-volume-fig-by-sex)

```{r sleep-outcomes-by-sex}
papaja::apa_table(
  model_tables_by_sex$sleep$data,
  caption = model_tables_by_sex$sleep$caption,
  note = model_tables_by_sex$sleep$note,
  col_spanners = model_tables_by_sex$sleep$col_spanners,
  escape = FALSE,
  landscape = TRUE
)
```

```{r sleep-by-volume-fig-by-sex, fig.cap = "Sleep metrics on Physical activity volume by sex"}
knitr::include_graphics(here::here(purdy_pictures_by_sex$predictor_pa_volume), error = FALSE)
```

### The effects of physical activity intensity on sleep by sex

We estimated how physical activity intensity affects sleep across sex.
We present the results in Table \@ref(tab:sleep-outcomes-by-sex) and Figure \@ref(fig:sleep-by-intensity-fig-by-sex).

```{r sleep-by-intensity-fig-by-sex, fig.cap = "Sleep metrics on Physical activity intensity moderated by sex"}
knitr::include_graphics(here::here(purdy_pictures_by_sex$predictor_pa_intensity), error = FALSE)
```

### The effects of sleep duration on physical activity

We estimated the effect of sleep duration on physical activity by sex.
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes-by-sex) and Figure \@ref(fig:PA-by-sleep-duration-fig-by-sex).

```{r pa-outcomes-by-sex}
papaja::apa_table(
  model_tables_by_sex$physical_activity$data,
  caption = model_tables_by_sex$physical_activity$caption,
  note = model_tables_by_sex$physical_activity$note,
  col_spanners = model_tables_by_sex$physical_activity$col_spanners,
  escape = FALSE,
  landscape = TRUE
)
```

```{r PA-by-sleep-duration-fig-by-sex, fig.cap = "Physical activity by sleep duration moderated by sex"}
knitr::include_graphics(here::here(purdy_pictures_by_sex$predictor_sleep_duration_lag), error = FALSE)
```

### The effects of sleep efficiency on physical activity

We estimated the effect of sleep efficiency on physical activity by sex.
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes-by-sex) and Figure \@ref(fig:PA-by-sleep-efficiency-fig-by-sex).

```{r PA-by-sleep-efficiency-fig-by-sex, fig.cap = "Physical activity by sleep efficiency moderated by sex"}
knitr::include_graphics(here::here(purdy_pictures_by_sex$predictor_sleep_efficiency_lag), error = FALSE)
```

### The effects of sleep onset on physical activity

We estimated the effect of sleep onset on physical activity by sex.
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes-by-sex) and Figure \@ref(fig:PA-by-sleep-onset-fig-by-sex).

```{r PA-by-sleep-onset-fig-by-sex, fig.cap = "Physical activity by sleep onset moderated by sex"}
knitr::include_graphics(here::here(purdy_pictures_by_sex$predictor_sleep_onset_lag), error = FALSE)
```

### The effects of sleep regularity on physical activity

We estimated the effect of sleep regularity on physical activity by sex.
Results are presented in Table \@ref(tab:pa-outcomes-by-sex) and Figure \@ref(fig:PA-by-sleep-regularity-fig-by-sex).

```{r PA-by-sleep-regularity-fig-by-sex, fig.cap = "Physical activity by sleep regularity moderated by sex"}
knitr::include_graphics(here::here(purdy_pictures_by_sex$predictor_sleep_regularity_lag), error = FALSE)
```
\clearpage

## weekday

```{r}
tar_load(model_tables_by_weekday)
tar_load(purdy_pictures_by_weekday)
```

### The effects of physical activity volume on sleep by weekday

We estimated the effects of physical activity on sleep (RQ1) using mixed-effects models.
We estimated the effect of physical activity volume on sleep by weekday, and the results are presented in Table \@ref(tab:sleep-outcomes-by-weekday) and Figure \@ref(fig:sleep-by-volume-fig-by-weekday)

```{r sleep-outcomes-by-weekday}
papaja::apa_table(
  model_tables_by_weekday$sleep$data,
  caption = model_tables_by_weekday$sleep$caption,
  note = model_tables_by_weekday$sleep$note,
  col_spanners = model_tables_by_weekday$sleep$col_spanners,
  escape = FALSE,
  landscape = TRUE
)
```

```{r sleep-by-volume-fig-by-weekday, fig.cap = "Sleep metrics on Physical activity volume by weekday"}
knitr::include_graphics(here::here(purdy_pictures_by_weekday$predictor_pa_volume), error = FALSE)
```

### The effects of physical activity intensity on sleep by weekday

We estimated how physical activity intensity affects sleep across weekday.
We present the results in Table \@ref(tab:sleep-outcomes-by-weekday) and Figure \@ref(fig:sleep-by-intensity-fig-by-weekday).

```{r sleep-by-intensity-fig-by-weekday, fig.cap = "Sleep metrics on Physical activity intensity moderated by weekday"}
knitr::include_graphics(here::here(purdy_pictures_by_weekday$predictor_pa_intensity), error = FALSE)
```

### The effects of sleep duration on physical activity

We estimated the effect of sleep duration on physical activity by weekday.
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes-by-weekday) and Figure \@ref(fig:PA-by-sleep-duration-fig-by-weekday).

```{r pa-outcomes-by-weekday}
papaja::apa_table(
  model_tables_by_weekday$physical_activity$data,
  caption = model_tables_by_weekday$physical_activity$caption,
  note = model_tables_by_weekday$physical_activity$note,
  col_spanners = model_tables_by_weekday$physical_activity$col_spanners,
  escape = FALSE,
  landscape = TRUE
)
```

```{r PA-by-sleep-duration-fig-by-weekday, fig.cap = "Physical activity by sleep duration moderated by weekday"}
knitr::include_graphics(here::here(purdy_pictures_by_weekday$predictor_sleep_duration_lag), error = FALSE)
```

### The effects of sleep efficiency on physical activity

We estimated the effect of sleep efficiency on physical activity by weekday.
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes-by-weekday) and Figure \@ref(fig:PA-by-sleep-efficiency-fig-by-weekday).

```{r PA-by-sleep-efficiency-fig-by-weekday, fig.cap = "Physical activity by sleep efficiency moderated by weekday"}
knitr::include_graphics(here::here(purdy_pictures_by_weekday$predictor_sleep_efficiency_lag), error = FALSE)
```

### The effects of sleep onset on physical activity

We estimated the effect of sleep onset on physical activity by weekday.
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes-by-weekday) and Figure \@ref(fig:PA-by-sleep-onset-fig-by-weekday).

```{r PA-by-sleep-onset-fig-by-weekday, fig.cap = "Physical activity by sleep onset moderated by weekday"}
knitr::include_graphics(here::here(purdy_pictures_by_weekday$predictor_sleep_onset_lag), error = FALSE)
```

### The effects of sleep regularity on physical activity

We estimated the effect of sleep regularity on physical activity by weekday.
Results are presented in Table \@ref(tab:pa-outcomes-by-weekday) and Figure \@ref(fig:PA-by-sleep-regularity-fig-by-weekday).

```{r PA-by-sleep-regularity-fig-by-weekday, fig.cap = "Physical activity by sleep regularity moderated by weekday"}
knitr::include_graphics(here::here(purdy_pictures_by_weekday$predictor_sleep_regularity_lag), error = FALSE)
```
\clearpage

## season

```{r}
tar_load(model_tables_by_season)
tar_load(purdy_pictures_by_season)
```

### The effects of physical activity volume on sleep by season

We estimated the effects of physical activity on sleep (RQ1) using mixed-effects models.
We estimated the effect of physical activity volume on sleep by season, and the results are presented in Table \@ref(tab:sleep-outcomes-by-season) and Figure \@ref(fig:sleep-by-volume-fig-by-season)

```{r sleep-outcomes-by-season}
papaja::apa_table(
  model_tables_by_season$sleep$data,
  caption = model_tables_by_season$sleep$caption,
  note = model_tables_by_season$sleep$note,
  col_spanners = model_tables_by_season$sleep$col_spanners,
  escape = FALSE,
  landscape = TRUE
)
```

```{r sleep-by-volume-fig-by-season, fig.cap = "Sleep metrics on Physical activity volume by season"}
knitr::include_graphics(here::here(purdy_pictures_by_season$predictor_pa_volume), error = FALSE)
```

### The effects of physical activity intensity on sleep by season

We estimated how physical activity intensity affects sleep across season.
We present the results in Table \@ref(tab:sleep-outcomes-by-season) and Figure \@ref(fig:sleep-by-intensity-fig-by-season).

```{r sleep-by-intensity-fig-by-season, fig.cap = "Sleep metrics on Physical activity intensity moderated by season"}
knitr::include_graphics(here::here(purdy_pictures_by_season$predictor_pa_intensity), error = FALSE)
```

### The effects of sleep duration on physical activity

We estimated the effect of sleep duration on physical activity by season.
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes-by-season) and Figure \@ref(fig:PA-by-sleep-duration-fig-by-season).

```{r pa-outcomes-by-season}
papaja::apa_table(
  model_tables_by_season$physical_activity$data,
  caption = model_tables_by_season$physical_activity$caption,
  note = model_tables_by_season$physical_activity$note,
  col_spanners = model_tables_by_season$physical_activity$col_spanners,
  escape = FALSE,
  landscape = TRUE
)
```

```{r PA-by-sleep-duration-fig-by-season, fig.cap = "Physical activity by sleep duration moderated by season"}
knitr::include_graphics(here::here(purdy_pictures_by_season$predictor_sleep_duration_lag), error = FALSE)
```

### The effects of sleep efficiency on physical activity

We estimated the effect of sleep efficiency on physical activity by season.
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes-by-season) and Figure \@ref(fig:PA-by-sleep-efficiency-fig-by-season).

```{r PA-by-sleep-efficiency-fig-by-season, fig.cap = "Physical activity by sleep efficiency moderated by season"}
knitr::include_graphics(here::here(purdy_pictures_by_season$predictor_sleep_efficiency_lag), error = FALSE)
```

### The effects of sleep onset on physical activity

We estimated the effect of sleep onset on physical activity by season.
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes-by-season) and Figure \@ref(fig:PA-by-sleep-onset-fig-by-season).

```{r PA-by-sleep-onset-fig-by-season, fig.cap = "Physical activity by sleep onset moderated by season"}
knitr::include_graphics(here::here(purdy_pictures_by_season$predictor_sleep_onset_lag), error = FALSE)
```

### The effects of sleep regularity on physical activity

We estimated the effect of sleep regularity on physical activity by season.
Results are presented in Table \@ref(tab:pa-outcomes-by-season) and Figure \@ref(fig:PA-by-sleep-regularity-fig-by-season).

```{r PA-by-sleep-regularity-fig-by-season, fig.cap = "Physical activity by sleep regularity moderated by season"}
knitr::include_graphics(here::here(purdy_pictures_by_season$predictor_sleep_regularity_lag), error = FALSE)
```
\clearpage

## region

```{r}
tar_load(model_tables_by_region)
tar_load(purdy_pictures_by_region)
```

### The effects of physical activity volume on sleep by region

We estimated the effects of physical activity on sleep (RQ1) using mixed-effects models.
We estimated the effect of physical activity volume on sleep by region, and the results are presented in Table \@ref(tab:sleep-outcomes-by-region) and Figure \@ref(fig:sleep-by-volume-fig-by-region)

```{r sleep-outcomes-by-region}
papaja::apa_table(
  model_tables_by_region$sleep$data,
  caption = model_tables_by_region$sleep$caption,
  note = model_tables_by_region$sleep$note,
  col_spanners = model_tables_by_region$sleep$col_spanners,
  escape = FALSE,
  landscape = TRUE
)
```

```{r sleep-by-volume-fig-by-region, fig.cap = "Sleep metrics on Physical activity volume by region"}
knitr::include_graphics(here::here(purdy_pictures_by_region$predictor_pa_volume), error = FALSE)
```

### The effects of physical activity intensity on sleep by region

We estimated how physical activity intensity affects sleep across region.
We present the results in Table \@ref(tab:sleep-outcomes-by-region) and Figure \@ref(fig:sleep-by-intensity-fig-by-region).

```{r sleep-by-intensity-fig-by-region, fig.cap = "Sleep metrics on Physical activity intensity moderated by region"}
knitr::include_graphics(here::here(purdy_pictures_by_region$predictor_pa_intensity), error = FALSE)
```

### The effects of sleep duration on physical activity

We estimated the effect of sleep duration on physical activity by region.
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes-by-region) and Figure \@ref(fig:PA-by-sleep-duration-fig-by-region).

```{r pa-outcomes-by-region}
papaja::apa_table(
  model_tables_by_region$physical_activity$data,
  caption = model_tables_by_region$physical_activity$caption,
  note = model_tables_by_region$physical_activity$note,
  col_spanners = model_tables_by_region$physical_activity$col_spanners,
  escape = FALSE,
  landscape = TRUE
)
```

```{r PA-by-sleep-duration-fig-by-region, fig.cap = "Physical activity by sleep duration moderated by region"}
knitr::include_graphics(here::here(purdy_pictures_by_region$predictor_sleep_duration_lag), error = FALSE)
```

### The effects of sleep efficiency on physical activity

We estimated the effect of sleep efficiency on physical activity by region.
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes-by-region) and Figure \@ref(fig:PA-by-sleep-efficiency-fig-by-region).

```{r PA-by-sleep-efficiency-fig-by-region, fig.cap = "Physical activity by sleep efficiency moderated by region"}
knitr::include_graphics(here::here(purdy_pictures_by_region$predictor_sleep_efficiency_lag), error = FALSE)
```

### The effects of sleep onset on physical activity

We estimated the effect of sleep onset on physical activity by region.
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes-by-region) and Figure \@ref(fig:PA-by-sleep-onset-fig-by-region).

```{r PA-by-sleep-onset-fig-by-region, fig.cap = "Physical activity by sleep onset moderated by region"}
knitr::include_graphics(here::here(purdy_pictures_by_region$predictor_sleep_onset_lag), error = FALSE)
```

### The effects of sleep regularity on physical activity

We estimated the effect of sleep regularity on physical activity by region.
Results are presented in Table \@ref(tab:pa-outcomes-by-region) and Figure \@ref(fig:PA-by-sleep-regularity-fig-by-region).

```{r PA-by-sleep-regularity-fig-by-region, fig.cap = "Physical activity by sleep regularity moderated by region"}
knitr::include_graphics(here::here(purdy_pictures_by_region$predictor_sleep_regularity_lag), error = FALSE)
```
\clearpage

## daylight hours

```{r}
tar_load(model_tables_by_daylight)
tar_load(purdy_pictures_by_daylight)
```

### The effects of physical activity volume on sleep by daylight hours

We estimated the effects of physical activity on sleep (RQ1) using mixed-effects models.
We estimated the effect of physical activity volume on sleep by daylight hours, and the results are presented in Table \@ref(tab:sleep-outcomes-by-daylight-hours) and Figure \@ref(fig:sleep-by-volume-fig-by-daylight-hours)

```{r sleep-outcomes-by-daylight-hours}
papaja::apa_table(
  model_tables_by_daylight$sleep$data,
  caption = model_tables_by_daylight$sleep$caption,
  note = model_tables_by_daylight$sleep$note,
  col_spanners = model_tables_by_daylight$sleep$col_spanners,
  escape = FALSE,
  landscape = TRUE
)
```

```{r sleep-by-volume-fig-by-daylight-hours, fig.cap = "Sleep metrics on Physical activity volume by daylight hours"}
knitr::include_graphics(here::here(purdy_pictures_by_daylight$predictor_pa_volume), error = FALSE)
```

### The effects of physical activity intensity on sleep by daylight hours

We estimated how physical activity intensity affects sleep across daylight hours.
We present the results in Table \@ref(tab:sleep-outcomes-by-daylight-hours) and Figure \@ref(fig:sleep-by-intensity-fig-by-daylight-hours).

```{r sleep-by-intensity-fig-by-daylight-hours, fig.cap = "Sleep metrics on Physical activity intensity moderated by daylight hours"}
knitr::include_graphics(here::here(purdy_pictures_by_daylight$predictor_pa_intensity), error = FALSE)
```

### The effects of sleep duration on physical activity

We estimated the effect of sleep duration on physical activity by daylight hours.
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes-by-daylight-hours) and Figure \@ref(fig:PA-by-sleep-duration-fig-by-daylight-hours).

```{r pa-outcomes-by-daylight-hours}
papaja::apa_table(
  model_tables_by_daylight$physical_activity$data,
  caption = model_tables_by_daylight$physical_activity$caption,
  note = model_tables_by_daylight$physical_activity$note,
  col_spanners = model_tables_by_daylight$physical_activity$col_spanners,
  escape = FALSE,
  landscape = TRUE
)
```

```{r PA-by-sleep-duration-fig-by-daylight-hours, fig.cap = "Physical activity by sleep duration moderated by daylight hours"}
knitr::include_graphics(here::here(purdy_pictures_by_daylight$predictor_sleep_duration_lag), error = FALSE)
```

### The effects of sleep efficiency on physical activity

We estimated the effect of sleep efficiency on physical activity by daylight hours.
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes-by-daylight-hours) and Figure \@ref(fig:PA-by-sleep-efficiency-fig-by-daylight-hours).

```{r PA-by-sleep-efficiency-fig-by-daylight-hours, fig.cap = "Physical activity by sleep efficiency moderated by daylight hours"}
knitr::include_graphics(here::here(purdy_pictures_by_daylight$predictor_sleep_efficiency_lag), error = FALSE)
```

### The effects of sleep onset on physical activity

We estimated the effect of sleep onset on physical activity by daylight hours.
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes-by-daylight-hours) and Figure \@ref(fig:PA-by-sleep-onset-fig-by-daylight-hours).

```{r PA-by-sleep-onset-fig-by-daylight-hours, fig.cap = "Physical activity by sleep onset moderated by daylight hours"}
knitr::include_graphics(here::here(purdy_pictures_by_daylight$predictor_sleep_onset_lag), error = FALSE)
```

### The effects of sleep regularity on physical activity

We estimated the effect of sleep regularity on physical activity by daylight hours.
Results are presented in Table \@ref(tab:pa-outcomes-by-daylight-hours) and Figure \@ref(fig:PA-by-sleep-regularity-fig-by-daylight-hours).

```{r PA-by-sleep-regularity-fig-by-daylight-hours, fig.cap = "Physical activity by sleep regularity moderated by daylight hours"}
knitr::include_graphics(here::here(purdy_pictures_by_daylight$predictor_sleep_regularity_lag), error = FALSE)
```
\clearpage

## wear location

```{r}
tar_load(model_tables_by_wear_location)
tar_load(purdy_pictures_by_wear_location)
```

### The effects of physical activity volume on sleep by wear location

We estimated the effects of physical activity on sleep (RQ1) using mixed-effects models.
We estimated the effect of physical activity volume on sleep by wear location, and the results are presented in Table \@ref(tab:sleep-outcomes-by-acc-wear-loc) and Figure \@ref(fig:sleep-by-volume-fig-by-acc-wear-loc)

```{r sleep-outcomes-by-acc-wear-loc}
papaja::apa_table(
  model_tables_by_wear_location$sleep$data,
  caption = model_tables_by_wear_location$sleep$caption,
  note = model_tables_by_wear_location$sleep$note,
  col_spanners = model_tables_by_wear_location$sleep$col_spanners,
  escape = FALSE,
  landscape = TRUE
)
```

```{r sleep-by-volume-fig-by-acc-wear-loc, fig.cap = "Sleep metrics on Physical activity volume by wear location"}
knitr::include_graphics(here::here(purdy_pictures_by_wear_location$predictor_pa_volume), error = FALSE)
```

### The effects of physical activity intensity on sleep by wear location

We estimated how physical activity intensity affects sleep across wear location.
We present the results in Table \@ref(tab:sleep-outcomes-by-acc-wear-loc) and Figure \@ref(fig:sleep-by-intensity-fig-by-acc-wear-loc).

```{r sleep-by-intensity-fig-by-acc-wear-loc, fig.cap = "Sleep metrics on Physical activity intensity moderated by wear location"}
knitr::include_graphics(here::here(purdy_pictures_by_wear_location$predictor_pa_intensity), error = FALSE)
```

### The effects of sleep duration on physical activity

We estimated the effect of sleep duration on physical activity by wear location.
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes-by-acc-wear-loc) and Figure \@ref(fig:PA-by-sleep-duration-fig-by-acc-wear-loc).

```{r pa-outcomes-by-acc-wear-loc}
papaja::apa_table(
  model_tables_by_wear_location$physical_activity$data,
  caption = model_tables_by_wear_location$physical_activity$caption,
  note = model_tables_by_wear_location$physical_activity$note,
  col_spanners = model_tables_by_wear_location$physical_activity$col_spanners,
  escape = FALSE,
  landscape = TRUE
)
```

```{r PA-by-sleep-duration-fig-by-acc-wear-loc, fig.cap = "Physical activity by sleep duration moderated by wear location"}
knitr::include_graphics(here::here(purdy_pictures_by_wear_location$predictor_sleep_duration_lag), error = FALSE)
```

### The effects of sleep efficiency on physical activity

We estimated the effect of sleep efficiency on physical activity by wear location.
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes-by-acc-wear-loc) and Figure \@ref(fig:PA-by-sleep-efficiency-fig-by-acc-wear-loc).

```{r PA-by-sleep-efficiency-fig-by-acc-wear-loc, fig.cap = "Physical activity by sleep efficiency moderated by wear location"}
knitr::include_graphics(here::here(purdy_pictures_by_wear_location$predictor_sleep_efficiency_lag), error = FALSE)
```

### The effects of sleep onset on physical activity

We estimated the effect of sleep onset on physical activity by wear location.
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes-by-acc-wear-loc) and Figure \@ref(fig:PA-by-sleep-onset-fig-by-acc-wear-loc).

```{r PA-by-sleep-onset-fig-by-acc-wear-loc, fig.cap = "Physical activity by sleep onset moderated by wear location"}
knitr::include_graphics(here::here(purdy_pictures_by_wear_location$predictor_sleep_onset_lag), error = FALSE)
```

### The effects of sleep regularity on physical activity

We estimated the effect of sleep regularity on physical activity by wear location.
Results are presented in Table \@ref(tab:pa-outcomes-by-acc-wear-loc) and Figure \@ref(fig:PA-by-sleep-regularity-fig-by-acc-wear-loc).

```{r PA-by-sleep-regularity-fig-by-acc-wear-loc, fig.cap = "Physical activity by sleep regularity moderated by wear location"}
knitr::include_graphics(here::here(purdy_pictures_by_wear_location$predictor_sleep_regularity_lag), error = FALSE)
```
\clearpage

## most active hour

```{r}
tar_load(model_tables_by_pa_mostactivehr)
tar_load(purdy_pictures_by_pa_mostactivehr)
```

### The effects of physical activity volume on sleep by most active hour

We estimated the effects of physical activity on sleep (RQ1) using mixed-effects models.
We estimated the effect of physical activity volume on sleep by most active hour, and the results are presented in Table \@ref(tab:sleep-outcomes-by-pa-mostactivehr) and Figure \@ref(fig:sleep-by-volume-fig-by-pa-mostactivehr)

```{r sleep-outcomes-by-pa-mostactivehr}
papaja::apa_table(
  model_tables_by_pa_mostactivehr$sleep$data,
  caption = model_tables_by_pa_mostactivehr$sleep$caption,
  note = model_tables_by_pa_mostactivehr$sleep$note,
  col_spanners = model_tables_by_pa_mostactivehr$sleep$col_spanners,
  escape = FALSE,
  landscape = TRUE
)
```

```{r sleep-by-volume-fig-by-pa-mostactivehr, fig.cap = "Sleep metrics on Physical activity volume by most active hour"}
knitr::include_graphics(here::here(purdy_pictures_by_pa_mostactivehr$predictor_pa_volume), error = FALSE)
```

### The effects of physical activity intensity on sleep by most active hour

We estimated how physical activity intensity affects sleep across most active hour.
We present the results in Table \@ref(tab:sleep-outcomes-by-pa-mostactivehr) and Figure \@ref(fig:sleep-by-intensity-fig-by-pa-mostactivehr).

```{r sleep-by-intensity-fig-by-pa-mostactivehr, fig.cap = "Sleep metrics on Physical activity intensity moderated by most active hour"}
knitr::include_graphics(here::here(purdy_pictures_by_pa_mostactivehr$predictor_pa_intensity), error = FALSE)
```

### The effects of sleep duration on physical activity

We estimated the effect of sleep duration on physical activity by most active hour.
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes-by-pa-mostactivehr) and Figure \@ref(fig:PA-by-sleep-duration-fig-by-pa-mostactivehr).

```{r pa-outcomes-by-pa-mostactivehr}
papaja::apa_table(
  model_tables_by_pa_mostactivehr$physical_activity$data,
  caption = model_tables_by_pa_mostactivehr$physical_activity$caption,
  note = model_tables_by_pa_mostactivehr$physical_activity$note,
  col_spanners = model_tables_by_pa_mostactivehr$physical_activity$col_spanners,
  escape = FALSE,
  landscape = TRUE
)
```

```{r PA-by-sleep-duration-fig-by-pa-mostactivehr, fig.cap = "Physical activity by sleep duration moderated by most active hour"}
knitr::include_graphics(here::here(purdy_pictures_by_pa_mostactivehr$predictor_sleep_duration_lag), error = FALSE)
```

### The effects of sleep efficiency on physical activity

We estimated the effect of sleep efficiency on physical activity by most active hour.
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes-by-pa-mostactivehr) and Figure \@ref(fig:PA-by-sleep-efficiency-fig-by-pa-mostactivehr).

```{r PA-by-sleep-efficiency-fig-by-pa-mostactivehr, fig.cap = "Physical activity by sleep efficiency moderated by most active hour"}
knitr::include_graphics(here::here(purdy_pictures_by_pa_mostactivehr$predictor_sleep_efficiency_lag), error = FALSE)
```

### The effects of sleep onset on physical activity

We estimated the effect of sleep onset on physical activity by most active hour.
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes-by-pa-mostactivehr) and Figure \@ref(fig:PA-by-sleep-onset-fig-by-pa-mostactivehr).

```{r PA-by-sleep-onset-fig-by-pa-mostactivehr, fig.cap = "Physical activity by sleep onset moderated by most active hour"}
knitr::include_graphics(here::here(purdy_pictures_by_pa_mostactivehr$predictor_sleep_onset_lag), error = FALSE)
```

### The effects of sleep regularity on physical activity

We estimated the effect of sleep regularity on physical activity by most active hour.
Results are presented in Table \@ref(tab:pa-outcomes-by-pa-mostactivehr) and Figure \@ref(fig:PA-by-sleep-regularity-fig-by-pa-mostactivehr).

```{r PA-by-sleep-regularity-fig-by-pa-mostactivehr, fig.cap = "Physical activity by sleep regularity moderated by most active hour"}
knitr::include_graphics(here::here(purdy_pictures_by_pa_mostactivehr$predictor_sleep_regularity_lag), error = FALSE)
```