### 2023-04-13 doc/multiverse_skeleton.Rmd
Move to static branching (#87)
* make_model_list takes pa and sleep vars as args

* Fix number of imputations

Fixes #86

* Add log_pa_volume variable

* Clear up linting message in data_imp

* Manage control vars dynamically

* Use static branching

* Add vars attribute to model_list

Co-authored-by: James Conigrave <12742211+conig@users.noreply.github.com>

* Add pipeline tweaks to allow for log variable

Co-authored-by: James Conigrave <12742211+conig@users.noreply.github.com>

* Rerun pipeline

* rename weekday_x

* Automate docs

Co-authored-by: James Conigrave <12742211+conig@users.norep

---
title             : "Multiverse analysis"
shorttitle        : "Title"

author: 
  - name          : "First Author"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "my@email.com"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - "Conceptualization"
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"
  - name          : "Ernst-August Doelle"
    affiliation   : "1,2"
    role:
      - "Writing - Review & Editing"
      - "Supervision"

affiliation:
  - id            : "1"
    institution   : "Wilhelm-Wundt-University"
  - id            : "2"
    institution   : "Konstanz Business School"

authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.

abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.

  <!-- https://tinyurl.com/ybremelq -->

keywords          : "keywords"
wordcount         : "X"

floatsintext      : no
linenumbers       : no
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no

classoption       : "man"

header-includes:
 - \usepackage[maxfloats=256]{morefloats}

output            : papaja::apa6_pdf

---

```{comment}
I needed to add maxdeadcycles and more floats to support how many figures are in this doc.
```

\renewcommand{\arraystretch}{0.7}
\maxdeadcycles=1000

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.pos = "!htbp")
source("R/utils.R")
options(warn = -1)
knitr::opts_chunk$set(out.width = "110%")
```

```{r load-libraries}
library(targets)
library(gt)
require(ggplot2)
library(papaja)
```

```{r load-targets, include=FALSE}
tar_load(multiverse_file)
pretty <- function(x) format(x, big.mark = ",")
```

# Results

## Model diagnostics

The model diagnostics for the main analysis, and for all models presented in this document are in Table \@ref(tab:model-diagnostics)

```{r model-diagnostics}
tar_read(model_diagnostics) |>
  papaja::apa_table(
    caption = "Model diagnostics", escape = FALSE,
    longtable = TRUE
  )
```

## Fixed Effect Model

```{r}
tar_load(model_tables_by_age_fixedef)
tar_load(purdy_pictures_by_age_fixedef)
```

In this document we show the results using another analytic technique.
Namely treating study ID as a fixed effect.
This is what we wrote in the protocol originally.

### The effects of physical activity volume on sleep

We estimated the effects of physical activity on sleep (RQ1) using fixed-effects models.
We estimated the effect of physical activity volume on sleep by age, and the results are presented in Table \@ref(tab:sleep-outcomes-fixedef) and Figure \@ref(fig:sleep-by-volume-fig-fixedef).

```{r sleep-outcomes-fixedef}
papaja::apa_table(
  model_tables_by_age_fixedef$sleep$data,
  caption = model_tables_by_age_fixedef$sleep$caption,
  note = model_tables_by_age_fixedef$sleep$note,
  col_spanners = model_tables_by_age_fixedef$sleep$col_spanners,
  escape = FALSE,
  landscape = TRUE
)
```

```{r sleep-by-volume-fig-fixedef, fig.cap = "Sleep metrics on Physical activity volume"}
knitr::include_graphics(here::here(purdy_pictures_by_age_fixedef$predictor_pa_volume), error = FALSE)
```

### The effects of physical activity intensity on sleep

We estimated how physical activity intensity affects sleep across different age groups.
We present the results controlling for sex, SES, and BMI, in Table \@ref(tab:sleep-outcomes-fixedef) and Figure \@ref(fig:sleep-by-intensity-fig-fixedef).

```{r sleep-by-intensity-fig-fixedef, fig.cap = "Sleep metrics on Physical activity intensity"}
knitr::include_graphics(here::here(purdy_pictures_by_age_fixedef$predictor_pa_intensity), error = FALSE)
```

### The effects of sleep duration on physical activity

We estimated the effect of sleep duration on physical activity by age.
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes-fixedef) and Figure \@ref(fig:PA-by-sleep-duration-fig-fixedef).

```{r pa-outcomes-fixedef}
papaja::apa_table(
  model_tables_by_age_fixedef$physical_activity$data,
  caption = model_tables_by_age_fixedef$physical_activity$caption,
  note = model_tables_by_age_fixedef$physical_activity$note,
  col_spanners = model_tables_by_age_fixedef$physical_activity$col_spanners,
  escape = FALSE,
  landscape = TRUE
)
```

```{r PA-by-sleep-duration-fig-fixedef, fig.cap = "Physical activity by sleep duration"}
knitr::include_graphics(here::here(purdy_pictures_by_age_fixedef$predictor_sleep_duration_lag), error = FALSE)
```

### The effects of sleep efficiency on physical activity

We estimated the effect of sleep efficiency on physical activity by age.
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes-fixedef) and Figure \@ref(fig:PA-by-sleep-efficiency-fig-fixedef).

```{r PA-by-sleep-efficiency-fig-fixedef, fig.cap = "Physical activity by sleep efficiency"}
knitr::include_graphics(here::here(purdy_pictures_by_age_fixedef$predictor_sleep_efficiency_lag), error = FALSE)
```

### The effects of sleep onset on physical activity

We estimated the effect of sleep onset on physical activity by age.
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes-fixedef) and Figure \@ref(fig:PA-by-sleep-onset-fig-fixedef).

```{r PA-by-sleep-onset-fig-fixedef, fig.cap = "Physical activity by sleep onset"}
knitr::include_graphics(here::here(purdy_pictures_by_age_fixedef$predictor_sleep_onset_lag), error = FALSE)
```

### The effects of sleep regularity on physical activity

We estimated the effect of sleep regularity on physical activity by age.
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes-fixedef) and Figure \@ref(fig:PA-by-sleep-regularity-fig-fixedef).

```{r PA-by-sleep-regularity-fig-fixedef, fig.cap = "Physical activity by sleep regularity"}
knitr::include_graphics(here::here(purdy_pictures_by_age_fixedef$predictor_sleep_regularity_lag), error = FALSE)
```

## Without Log Transforming PA Volume

```{r}
tar_load(model_tables_by_age_nolog)
tar_load(purdy_pictures_by_age_nolog)
```

Based on preliminary modelling on a subset of the data, we chose to log transform the physical activity volume variable.
Below, we present the results had we not made this decision.
These results are all moderated by age, as this was our primary analysis.

### The effects of physical activity volume on sleep

We estimated the effects of physical activity on sleep (RQ1) using mixed-effects models.
We estimated the effect of physical activity volume on sleep by age, and the results are presented in Table \@ref(tab:sleep-outcomes-nolog) and Figure \@ref(fig:sleep-by-volume-fig-nolog).

```{r sleep-outcomes-nolog}
papaja::apa_table(
  model_tables_by_age_nolog$sleep$data,
  caption = model_tables_by_age_nolog$sleep$caption,
  note = model_tables_by_age_nolog$sleep$note,
  col_spanners = model_tables_by_age_nolog$sleep$col_spanners,
  escape = FALSE,
  landscape = TRUE
)
```

```{r sleep-by-volume-fig-nolog, fig.cap = "Sleep metrics on Physical activity volume"}
knitr::include_graphics(here::here(purdy_pictures_by_age_nolog$predictor_pa_volume), error = FALSE)
```

### The effects of sleep duration on physical activity

We estimated the effect of sleep duration on physical activity by age.
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes-nolog) and Figure \@ref(fig:PA-by-sleep-duration-fig-nolog).

```{r pa-outcomes-nolog}
papaja::apa_table(
  model_tables_by_age_nolog$physical_activity$data,
  caption = model_tables_by_age_nolog$physical_activity$caption,
  note = model_tables_by_age_nolog$physical_activity$note,
  col_spanners = model_tables_by_age_nolog$physical_activity$col_spanners,
  escape = FALSE,
  landscape = TRUE
)
```

```{r PA-by-sleep-duration-fig-nolog, fig.cap = "Physical activity by sleep duration"}
knitr::include_graphics(here::here(purdy_pictures_by_age_nolog$predictor_sleep_duration_lag), error = FALSE)
```

### The effects of sleep efficiency on physical activity

We estimated the effect of sleep efficiency on physical activity by age.
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes-nolog) and Figure \@ref(fig:PA-by-sleep-efficiency-fig-nolog).

```{r PA-by-sleep-efficiency-fig-nolog, fig.cap = "Physical activity by sleep efficiency"}
knitr::include_graphics(here::here(purdy_pictures_by_age_nolog$predictor_sleep_efficiency_lag), error = FALSE)
```

### The effects of sleep onset on physical activity

We estimated the effect of sleep onset on physical activity by age.
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes-nolog) and Figure \@ref(fig:PA-by-sleep-onset-fig-nolog).

```{r PA-by-sleep-onset-fig-nolog, fig.cap = "Physical activity by sleep onset"}
knitr::include_graphics(here::here(purdy_pictures_by_age_nolog$predictor_sleep_onset_lag), error = FALSE)
```

### The effects of sleep regularity on physical activity

We estimated the effect of sleep regularity on physical activity by age.
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes-nolog) and Figure \@ref(fig:PA-by-sleep-regularity-fig-nolog).

```{r PA-by-sleep-regularity-fig-nolog, fig.cap = "Physical activity by sleep regularity"}
knitr::include_graphics(here::here(purdy_pictures_by_age_nolog$predictor_sleep_regularity_lag), error = FALSE)
```

# Moderation

In the following section we present models for a range of moderators which were pre-defined in our protocol.