### 2023-04-21 doc/multiverse_skeleton.Rmd
Generate presentation names for tables. Back transform figs (#90)
* Generate presentation names for tables.

* Send scale var descriptives down with the model_list for exact transformation.

* Back transform figures.

* Increase range of variables to facilitate back-transformed variables.

* Fix kurtosis, set range for tile plots.

* Change PA-volume back to unlogged.

* Clean up empty function.

---
title             : "Multiverse analysis"
shorttitle        : "Title"

author: 
  - name          : "First Author"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "my@email.com"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - "Conceptualization"
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"
  - name          : "Ernst-August Doelle"
    affiliation   : "1,2"
    role:
      - "Writing - Review & Editing"
      - "Supervision"

affiliation:
  - id            : "1"
    institution   : "Wilhelm-Wundt-University"
  - id            : "2"
    institution   : "Konstanz Business School"

authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.

abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.

  <!-- https://tinyurl.com/ybremelq -->

keywords          : "keywords"
wordcount         : "X"

floatsintext      : no
linenumbers       : no
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no

classoption       : "man"

header-includes:
 - \usepackage[maxfloats=256]{morefloats}

output            : papaja::apa6_pdf

---

```{comment}
I needed to add maxdeadcycles and more floats to support how many figures are in this doc.
```

\renewcommand{\arraystretch}{0.7}
\maxdeadcycles=1000

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.pos = "!htbp")
source("R/utils.R")
options(warn = -1)
knitr::opts_chunk$set(out.width = "110%")
```

```{r load-libraries}
library(targets)
library(gt)
require(ggplot2)
library(papaja)
```

```{r load-targets, include=FALSE}
tar_load(multiverse_file)
pretty <- function(x) format(x, big.mark = ",")
```

# Results

## Model diagnostics

The model diagnostics for the main analysis, and for all models presented in this document are in Table \@ref(tab:model-diagnostics)

```{r model-diagnostics}
tar_read(model_diagnostics) |>
  papaja::apa_table(
    caption = "Model diagnostics", escape = FALSE,
    longtable = TRUE
  )
```

## Fixed Effect Model

```{r}
tar_load(model_tables_by_age_fixedef)
tar_load(purdy_pictures_by_age_fixedef)
```

In this document we show the results using another analytic technique.
Namely treating study ID as a fixed effect.
This is what we wrote in the protocol originally.

### The effects of physical activity volume on sleep

We estimated the effects of physical activity on sleep (RQ1) using fixed-effects models.
We estimated the effect of physical activity volume on sleep by age, and the results are presented in Table \@ref(tab:sleep-outcomes-fixedef) and Figure \@ref(fig:sleep-by-volume-fig-fixedef).

```{r sleep-outcomes-fixedef}
papaja::apa_table(
  model_tables_by_age_fixedef$sleep$data,
  caption = model_tables_by_age_fixedef$sleep$caption,
  note = model_tables_by_age_fixedef$sleep$note,
  col_spanners = model_tables_by_age_fixedef$sleep$col_spanners,
  escape = FALSE,
  landscape = TRUE
)
```

```{r sleep-by-volume-fig-fixedef, fig.cap = "Sleep metrics on Physical activity volume"}
knitr::include_graphics(here::here(purdy_pictures_by_age_fixedef$predictor_pa_volume), error = FALSE)
```

### The effects of physical activity intensity on sleep

We estimated how physical activity intensity affects sleep across different age groups.
We present the results controlling for sex, SES, and BMI, in Table \@ref(tab:sleep-outcomes-fixedef) and Figure \@ref(fig:sleep-by-intensity-fig-fixedef).

```{r sleep-by-intensity-fig-fixedef, fig.cap = "Sleep metrics on Physical activity intensity"}
knitr::include_graphics(here::here(purdy_pictures_by_age_fixedef$predictor_pa_intensity), error = FALSE)
```

### The effects of sleep duration on physical activity

We estimated the effect of sleep duration on physical activity by age.
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes-fixedef) and Figure \@ref(fig:PA-by-sleep-duration-fig-fixedef).

```{r pa-outcomes-fixedef}
papaja::apa_table(
  model_tables_by_age_fixedef$physical_activity$data,
  caption = model_tables_by_age_fixedef$physical_activity$caption,
  note = model_tables_by_age_fixedef$physical_activity$note,
  col_spanners = model_tables_by_age_fixedef$physical_activity$col_spanners,
  escape = FALSE,
  landscape = TRUE
)
```

```{r PA-by-sleep-duration-fig-fixedef, fig.cap = "Physical activity by sleep duration"}
knitr::include_graphics(here::here(purdy_pictures_by_age_fixedef$predictor_sleep_duration_lag), error = FALSE)
```

### The effects of sleep efficiency on physical activity

We estimated the effect of sleep efficiency on physical activity by age.
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes-fixedef) and Figure \@ref(fig:PA-by-sleep-efficiency-fig-fixedef).

```{r PA-by-sleep-efficiency-fig-fixedef, fig.cap = "Physical activity by sleep efficiency"}
knitr::include_graphics(here::here(purdy_pictures_by_age_fixedef$predictor_sleep_efficiency_lag), error = FALSE)
```

### The effects of sleep onset on physical activity

We estimated the effect of sleep onset on physical activity by age.
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes-fixedef) and Figure \@ref(fig:PA-by-sleep-onset-fig-fixedef).

```{r PA-by-sleep-onset-fig-fixedef, fig.cap = "Physical activity by sleep onset"}
knitr::include_graphics(here::here(purdy_pictures_by_age_fixedef$predictor_sleep_onset_lag), error = FALSE)
```

### The effects of sleep regularity on physical activity

We estimated the effect of sleep regularity on physical activity by age.
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes-fixedef) and Figure \@ref(fig:PA-by-sleep-regularity-fig-fixedef).

```{r PA-by-sleep-regularity-fig-fixedef, fig.cap = "Physical activity by sleep regularity"}
knitr::include_graphics(here::here(purdy_pictures_by_age_fixedef$predictor_sleep_regularity_lag), error = FALSE)
```

## Log Transforming PA Volume

```{r}
tar_load(model_tables_by_age_log)
tar_load(purdy_pictures_by_age_log)
```

Based on preliminary modelling on a subset of the data, we chose to leave physical activity volume in its natural form to enhance interpretability. The log scale meant that the variable became unintepretable towards the edge of its range. Below, we present the results had we not made this decision. These results are all moderated by age, as this was our primary analysis. The associated figures are back-transformed to the scale range.

### The effects of physical activity volume on sleep

We estimated the effects of physical activity on sleep (RQ1) using mixed-effects models.
We estimated the effect of physical activity volume on sleep by age, and the results are presented in Table \@ref(tab:sleep-outcomes-log) and Figure \@ref(fig:sleep-by-volume-fig-log).

```{r sleep-outcomes-log}
papaja::apa_table(
  model_tables_by_age_log$sleep$data,
  caption = model_tables_by_age_log$sleep$caption,
  note = model_tables_by_age_log$sleep$note,
  col_spanners = model_tables_by_age_log$sleep$col_spanners,
  escape = FALSE,
  landscape = TRUE
)
```

```{r sleep-by-volume-fig-log, fig.cap = "Sleep metrics on Physical activity volume"}
knitr::include_graphics(here::here(purdy_pictures_by_age_log$predictor_pa_volume), error = FALSE)
```

### The effects of sleep duration on physical activity

We estimated the effect of sleep duration on physical activity by age.
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes-log) and Figure \@ref(fig:PA-by-sleep-duration-fig-log).

```{r pa-outcomes-log}
papaja::apa_table(
  model_tables_by_age_log$physical_activity$data,
  caption = model_tables_by_age_log$physical_activity$caption,
  note = model_tables_by_age_log$physical_activity$note,
  col_spanners = model_tables_by_age_log$physical_activity$col_spanners,
  escape = FALSE,
  landscape = TRUE
)
```

```{r PA-by-sleep-duration-fig-log, fig.cap = "Physical activity by sleep duration"}
knitr::include_graphics(here::here(purdy_pictures_by_age_log$predictor_sleep_duration_lag), error = FALSE)
```

### The effects of sleep efficiency on physical activity

We estimated the effect of sleep efficiency on physical activity by age.
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes-log) and Figure \@ref(fig:PA-by-sleep-efficiency-fig-log).

```{r PA-by-sleep-efficiency-fig-log, fig.cap = "Physical activity by sleep efficiency"}
knitr::include_graphics(here::here(purdy_pictures_by_age_log$predictor_sleep_efficiency_lag), error = FALSE)
```

### The effects of sleep onset on physical activity

We estimated the effect of sleep onset on physical activity by age.
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes-log) and Figure \@ref(fig:PA-by-sleep-onset-fig-log).

```{r PA-by-sleep-onset-fig-log, fig.cap = "Physical activity by sleep onset"}
knitr::include_graphics(here::here(purdy_pictures_by_age_log$predictor_sleep_onset_lag), error = FALSE)
```

### The effects of sleep regularity on physical activity

We estimated the effect of sleep regularity on physical activity by age.
Results, controlling for sex, SES, and BMI are presented in Table \@ref(tab:pa-outcomes-log) and Figure \@ref(fig:PA-by-sleep-regularity-fig-log).

```{r PA-by-sleep-regularity-fig-log, fig.cap = "Physical activity by sleep regularity"}
knitr::include_graphics(here::here(purdy_pictures_by_age_log$predictor_sleep_regularity_lag), error = FALSE)
```

# Moderation

In the following section we present models for a range of moderators which were pre-defined in our protocol.